---
title: "Time dependent AUC with real data_extra"
author: "Seong-Yun Hwang"
date: "2023-04-04"
output: html_document
---

# Loading packages

```{r warning = FALSE, message = FALSE}
library(survival) 
library(timeROC)
library(timereg)
library(ggplot2)
library(reshape)
library(plyr)
library(tidyverse)
library(coxed)
library(gridExtra)
library(mgcv)
```


# Real survival data (in package 'survival')

```{r}
data(pbc)
head(pbc)
pbc$status[(pbc$status == 1)] = 0
pbc$status[(pbc$status == 2)] = 1
pbc$edema <- as.factor(pbc$edema)
```



# 3 explanatory variables

## Time dependent AUC (Cumulative/Dynamic time-dependent ROC curve)

```{r}
dt5 <- na.omit(pbc)
cox.dep.dt5 <- coxph(Surv(time,status) ~ edema, data=dt5)
summary(cox.dep.dt5)
```

```{r}
risk.dep.dt5 <- predict(object=cox.dep.dt5, newdata=dt5, type="risk") # risk score
dt5$risk.dep.dt5 <- risk.dep.dt5
ROC.risk.dep.dt5 <- timeROC(T=dt5$time,
                         delta=dt5$status,marker=dt5$risk.dep.dt5,
                         cause=1,weighting="marginal",
                         times=dt5$time)
ROC.risk.dep.dt5
```

```{r}
TP1 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1191"][,1]
FP1 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1191"][,1]
TP2 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1786"]
FP2 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1786"]
TP3 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=2689"]
FP3 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=2689"]
TP4 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=3422"]
FP4 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=3422"]
TP5 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=4032"]
FP5 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=4032"]
TP <- c(TP1,TP2,TP3,TP4,TP5)
FP <- c(FP1,FP2,FP3,FP4,FP5)
time.survival <- c(rep("t=1191",4),rep("t=1786",4),rep("t=2689",4),
                   rep("t=3422",4),rep("t=4032",4))
dt6 <- data.frame(times=time.survival,TP=TP,FP=FP)
```

```{r}
ggplot(data=dt6)+
  geom_line(mapping=aes(x=FP,
                        y=TP,
                        group=times,
                        color=times),
            size=1) +
  ggtitle("Time dependent ROC curve (pbc data) - 1 explanatory variable") +
  xlab("False Positive Rate") + ylab("True Positive Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```



## Make AUC data

```{r}
# uncensored
uncensored.density5 <- density(dt5$time[dt5$status==1])
data5_1 <- data.frame(time=uncensored.density5[["x"]], 
                    density=uncensored.density5[["y"]])
data5_1$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(uncensored.density5[["y"]]-min(uncensored.density5[["y"]]))/(max(uncensored.density5[["y"]])-min(uncensored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

# censored
censored.density5 <- density(dt5$time[dt5$status==0])
data5_2 <- data.frame(time=censored.density5[["x"]], 
                    density=censored.density5[["y"]])
data5_2$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(censored.density5[["y"]]-min(censored.density5[["y"]]))/(max(censored.density5[["y"]])-min(censored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

res5 <- data.frame(time=ROC.risk.dep.dt5[["times"]],
                    cox.dep.dt5=ROC.risk.dep.dt5[["AUC"]])
res5 <- melt(data = res5, 
                   id.vars = "time", 
                   measure.vars = c("cox.dep.dt5"))
```


## Time dependent AUC plot with ggplot2 

```{r}
dt5$censor1 <- ifelse(dt5$status==1,
                    1,min(na.omit(ROC.risk.dep.dt5[["AUC"]])))
ggplot() +
  geom_line(data=res5, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt5, aes(x=time, y=censor1, color=as.factor(status))) +
  geom_line(data=data5_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data5_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, max(dt5$time))) +
  geom_vline(xintercept = min(dt5$time[dt5$status==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Time dependent AUC with risk score (pbc data) - 1 explanatory variable") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```









# 3 explanatory variables

## Time dependent AUC (Cumulative/Dynamic time-dependent ROC curve)

```{r}
dt5 <- na.omit(pbc)
cox.dep.dt5 <- coxph(Surv(time,status) ~ edema+bili+albumin, data=dt5)
summary(cox.dep.dt5)
```

```{r}
risk.dep.dt5 <- predict(object=cox.dep.dt5, newdata=dt5, type="risk") # risk score
dt5$risk.dep.dt5 <- risk.dep.dt5
ROC.risk.dep.dt5 <- timeROC(T=dt5$time,
                         delta=dt5$status,marker=dt5$risk.dep.dt5,
                         cause=1,weighting="marginal",
                         times=dt5$time)
ROC.risk.dep.dt5
```

```{r}
TP1 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1191"][,1]
FP1 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1191"][,1]
TP2 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1786"]
FP2 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1786"]
TP3 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=2689"]
FP3 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=2689"]
TP4 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=3422"]
FP4 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=3422"]
TP5 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=4032"]
FP5 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=4032"]
TP <- c(TP1,TP2,TP3,TP4,TP5)
FP <- c(FP1,FP2,FP3,FP4,FP5)
time.survival <- c(rep("t=1191",272),rep("t=1786",272),rep("t=2689",272),
                   rep("t=3422",272),rep("t=4032",272))
dt6 <- data.frame(times=time.survival,TP=TP,FP=FP)
```

```{r}
ggplot(data=dt6)+
  geom_line(mapping=aes(x=FP,
                        y=TP,
                        group=times,
                        color=times),
            size=1) +
  ggtitle("Time dependent ROC curve (pbc data) - 3 explanatory variables") +
  xlab("False Positive Rate") + ylab("True Positive Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```



## Make AUC data

```{r}
# uncensored
uncensored.density5 <- density(dt5$time[dt5$status==1])
data5_1 <- data.frame(time=uncensored.density5[["x"]], 
                    density=uncensored.density5[["y"]])
data5_1$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(uncensored.density5[["y"]]-min(uncensored.density5[["y"]]))/(max(uncensored.density5[["y"]])-min(uncensored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

# censored
censored.density5 <- density(dt5$time[dt5$status==0])
data5_2 <- data.frame(time=censored.density5[["x"]], 
                    density=censored.density5[["y"]])
data5_2$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(censored.density5[["y"]]-min(censored.density5[["y"]]))/(max(censored.density5[["y"]])-min(censored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

res5 <- data.frame(time=ROC.risk.dep.dt5[["times"]],
                    cox.dep.dt5=ROC.risk.dep.dt5[["AUC"]])
res5 <- melt(data = res5, 
                   id.vars = "time", 
                   measure.vars = c("cox.dep.dt5"))
```


## Time dependent AUC plot with ggplot2 

```{r}
dt5$censor1 <- ifelse(dt5$status==1,
                    1,min(na.omit(ROC.risk.dep.dt5[["AUC"]])))
ggplot() +
  geom_line(data=res5, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt5, aes(x=time, y=censor1, color=as.factor(status))) +
  geom_line(data=data5_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data5_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, max(dt5$time))) +
  geom_vline(xintercept = min(dt5$time[dt5$status==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Time dependent AUC with risk score (pbc data) - 3 explanatory variables") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```








# 5 explanatory variables

## Time dependent AUC (Cumulative/Dynamic time-dependent ROC curve)

```{r}
dt5 <- na.omit(pbc)
cox.dep.dt5 <- coxph(Surv(time,status) ~ edema+bili+albumin+age+copper, data=dt5)
summary(cox.dep.dt5)
```

```{r}
risk.dep.dt5 <- predict(object=cox.dep.dt5, newdata=dt5, type="risk") # risk score
dt5$risk.dep.dt5 <- risk.dep.dt5
ROC.risk.dep.dt5 <- timeROC(T=dt5$time,
                         delta=dt5$status,marker=dt5$risk.dep.dt5,
                         cause=1,weighting="marginal",
                         times=dt5$time)
ROC.risk.dep.dt5
```

```{r}
TP1 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1191"][,1]
FP1 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1191"][,1]
TP2 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1786"]
FP2 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1786"]
TP3 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=2689"]
FP3 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=2689"]
TP4 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=3422"]
FP4 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=3422"]
TP5 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=4032"]
FP5 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=4032"]
TP <- c(TP1,TP2,TP3,TP4,TP5)
FP <- c(FP1,FP2,FP3,FP4,FP5)
time.survival <- c(rep("t=1191",277),rep("t=1786",277),rep("t=2689",277),
                   rep("t=3422",277),rep("t=4032",277))
dt6 <- data.frame(times=time.survival,TP=TP,FP=FP)
```

```{r}
ggplot(data=dt6)+
  geom_line(mapping=aes(x=FP,
                        y=TP,
                        group=times,
                        color=times),
            size=1) +
  ggtitle("Time dependent ROC curve (pbc data) - 5 explanatory variables") +
  xlab("False Positive Rate") + ylab("True Positive Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```



## Make AUC data

```{r}
# uncensored
uncensored.density5 <- density(dt5$time[dt5$status==1])
data5_1 <- data.frame(time=uncensored.density5[["x"]], 
                    density=uncensored.density5[["y"]])
data5_1$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(uncensored.density5[["y"]]-min(uncensored.density5[["y"]]))/(max(uncensored.density5[["y"]])-min(uncensored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

# censored
censored.density5 <- density(dt5$time[dt5$status==0])
data5_2 <- data.frame(time=censored.density5[["x"]], 
                    density=censored.density5[["y"]])
data5_2$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(censored.density5[["y"]]-min(censored.density5[["y"]]))/(max(censored.density5[["y"]])-min(censored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

res5 <- data.frame(time=ROC.risk.dep.dt5[["times"]],
                    cox.dep.dt5=ROC.risk.dep.dt5[["AUC"]])
res5 <- melt(data = res5, 
                   id.vars = "time", 
                   measure.vars = c("cox.dep.dt5"))
```


## Time dependent AUC plot with ggplot2 

```{r}
dt5$censor1 <- ifelse(dt5$status==1,
                    1,min(na.omit(ROC.risk.dep.dt5[["AUC"]])))
ggplot() +
  geom_line(data=res5, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt5, aes(x=time, y=censor1, color=as.factor(status))) +
  geom_line(data=data5_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data5_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, max(dt5$time))) +
  geom_vline(xintercept = min(dt5$time[dt5$status==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Time dependent AUC with risk score (pbc data) - 5 explanatory variables") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```








# 7 explanatory variables

## Time dependent AUC (Cumulative/Dynamic time-dependent ROC curve)

```{r}
dt5 <- na.omit(pbc)
cox.dep.dt5 <- coxph(Surv(time,status) ~ edema+bili+albumin+age+copper+protime+stage, data=dt5)
summary(cox.dep.dt5)
```

```{r}
risk.dep.dt5 <- predict(object=cox.dep.dt5, newdata=dt5, type="risk") # risk score
dt5$risk.dep.dt5 <- risk.dep.dt5
ROC.risk.dep.dt5 <- timeROC(T=dt5$time,
                         delta=dt5$status,marker=dt5$risk.dep.dt5,
                         cause=1,weighting="marginal",
                         times=dt5$time)
ROC.risk.dep.dt5
```

```{r}
TP1 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1191"][,1]
FP1 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1191"][,1]
TP2 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1786"]
FP2 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1786"]
TP3 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=2689"]
FP3 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=2689"]
TP4 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=3422"]
FP4 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=3422"]
TP5 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=4032"]
FP5 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=4032"]
TP <- c(TP1,TP2,TP3,TP4,TP5)
FP <- c(FP1,FP2,FP3,FP4,FP5)
time.survival <- c(rep("t=1191",277),rep("t=1786",277),rep("t=2689",277),
                   rep("t=3422",277),rep("t=4032",277))
dt6 <- data.frame(times=time.survival,TP=TP,FP=FP)
```

```{r}
ggplot(data=dt6)+
  geom_line(mapping=aes(x=FP,
                        y=TP,
                        group=times,
                        color=times),
            size=1) +
  ggtitle("Time dependent ROC curve (pbc data) - 7 explanatory variables") +
  xlab("False Positive Rate") + ylab("True Positive Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```



## Make AUC data

```{r}
# uncensored
uncensored.density5 <- density(dt5$time[dt5$status==1])
data5_1 <- data.frame(time=uncensored.density5[["x"]], 
                    density=uncensored.density5[["y"]])
data5_1$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(uncensored.density5[["y"]]-min(uncensored.density5[["y"]]))/(max(uncensored.density5[["y"]])-min(uncensored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

# censored
censored.density5 <- density(dt5$time[dt5$status==0])
data5_2 <- data.frame(time=censored.density5[["x"]], 
                    density=censored.density5[["y"]])
data5_2$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(censored.density5[["y"]]-min(censored.density5[["y"]]))/(max(censored.density5[["y"]])-min(censored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

res5 <- data.frame(time=ROC.risk.dep.dt5[["times"]],
                    cox.dep.dt5=ROC.risk.dep.dt5[["AUC"]])
res5 <- melt(data = res5, 
                   id.vars = "time", 
                   measure.vars = c("cox.dep.dt5"))
```


## Time dependent AUC plot with ggplot2 

```{r}
dt5$censor1 <- ifelse(dt5$status==1,
                    1,min(na.omit(ROC.risk.dep.dt5[["AUC"]])))
ggplot() +
  geom_line(data=res5, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt5, aes(x=time, y=censor1, color=as.factor(status))) +
  geom_line(data=data5_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data5_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, max(dt5$time))) +
  geom_vline(xintercept = min(dt5$time[dt5$status==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Time dependent AUC with risk score (pbc data) - 7 explanatory variables") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```








# 9 explanatory variables

## Time dependent AUC (Cumulative/Dynamic time-dependent ROC curve)

```{r}
dt5 <- na.omit(pbc)
cox.dep.dt5 <- coxph(Surv(time,status) ~ edema+bili+albumin+age+copper+protime+stage+ast+sex, data=dt5)
summary(cox.dep.dt5)
```

```{r}
risk.dep.dt5 <- predict(object=cox.dep.dt5, newdata=dt5, type="risk") # risk score
dt5$risk.dep.dt5 <- risk.dep.dt5
ROC.risk.dep.dt5 <- timeROC(T=dt5$time,
                         delta=dt5$status,marker=dt5$risk.dep.dt5,
                         cause=1,weighting="marginal",
                         times=dt5$time)
ROC.risk.dep.dt5
```

```{r}
TP1 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1191"][,1]
FP1 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1191"][,1]
TP2 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=1786"]
FP2 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=1786"]
TP3 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=2689"]
FP3 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=2689"]
TP4 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=3422"]
FP4 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=3422"]
TP5 <- ROC.risk.dep.dt5[["TP"]][,colnames(ROC.risk.dep.dt5[["TP"]])=="t=4032"]
FP5 <- ROC.risk.dep.dt5[["FP"]][,colnames(ROC.risk.dep.dt5[["FP"]])=="t=4032"]
TP <- c(TP1,TP2,TP3,TP4,TP5)
FP <- c(FP1,FP2,FP3,FP4,FP5)
time.survival <- c(rep("t=1191",277),rep("t=1786",277),rep("t=2689",277),
                   rep("t=3422",277),rep("t=4032",277))
dt6 <- data.frame(times=time.survival,TP=TP,FP=FP)
```

```{r}
ggplot(data=dt6)+
  geom_line(mapping=aes(x=FP,
                        y=TP,
                        group=times,
                        color=times),
            size=1) +
  ggtitle("Time dependent ROC curve (pbc data) - 9 explanatory variables") +
  xlab("False Positive Rate") + ylab("True Positive Rate") +
  theme(plot.title = element_text(hjust = 0.5))
```



## Make AUC data

```{r}
# uncensored
uncensored.density5 <- density(dt5$time[dt5$status==1])
data5_1 <- data.frame(time=uncensored.density5[["x"]], 
                    density=uncensored.density5[["y"]])
data5_1$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(uncensored.density5[["y"]]-min(uncensored.density5[["y"]]))/(max(uncensored.density5[["y"]])-min(uncensored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

# censored
censored.density5 <- density(dt5$time[dt5$status==0])
data5_2 <- data.frame(time=censored.density5[["x"]], 
                    density=censored.density5[["y"]])
data5_2$density.adjusted <- (max(na.omit(ROC.risk.dep.dt5[["AUC"]]))-min(na.omit(ROC.risk.dep.dt5[["AUC"]])))*(censored.density5[["y"]]-min(censored.density5[["y"]]))/(max(censored.density5[["y"]])-min(censored.density5[["y"]]))+min(na.omit(ROC.risk.dep.dt5[["AUC"]]))

res5 <- data.frame(time=ROC.risk.dep.dt5[["times"]],
                    cox.dep.dt5=ROC.risk.dep.dt5[["AUC"]])
res5 <- melt(data = res5, 
                   id.vars = "time", 
                   measure.vars = c("cox.dep.dt5"))
```


## Time dependent AUC plot with ggplot2 

```{r}
dt5$censor1 <- ifelse(dt5$status==1,
                    1,min(na.omit(ROC.risk.dep.dt5[["AUC"]])))
ggplot() +
  geom_line(data=res5, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt5, aes(x=time, y=censor1, color=as.factor(status))) +
  geom_line(data=data5_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data5_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, max(dt5$time))) +
  geom_vline(xintercept = min(dt5$time[dt5$status==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Time dependent AUC with risk score (pbc data) - 9 explanatory variables") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

