---
title: "Survival tree in R"
author: "Hwang Seong-Yun"
date: '2022 3 3 '
output: html_document
---

# Usage of R package LTRCtrees

- Reference : https://cran.r-project.org/web/packages/LTRCtrees/vignettes/LTRCtrees.html

## Example of fitting survival trees for LTRC data

### The assay of serum free light chain data example

```{r}
## Adjust data & clean data
library(survival)
set.seed(0)
## Since LTRCART uses cross-validation to prune the tree, specifying the seed 
## guarantees that the results given here will be duplicated in other analyses
Data <- flchain
Data <- Data[!is.na(Data$creatinine),]
Data$End <- Data$age + Data$futime/365
DATA <- Data[Data$End > Data$age,]
names(DATA)[6] <- "FLC"
head(DATA)
```

```{r}
## Setup training set and test set
Train = DATA[1:500,]
Test = DATA[1000:1020,]
```

```{r}
## Fit LTRCART and LTRCIT survival tree
library(LTRCtrees)
LTRCART.obj <- LTRCART(Surv(age, End, death) ~ sex + FLC + creatinine, Train) # Fit a relative risk survival tree for LTRC(left-truncated and right censored) data
LTRCIT.obj <- LTRCIT(Surv(age, End, death) ~ sex + FLC + creatinine, Train) # Fit a conditional inference survival tree for LTRC data
```

```{r}
## Putting Surv(End, death) in formula would result an error message
## since both LTRCART and LTRCIT are expecting Surv(time1, time2, event)
 
## Plot the fitted LTRCART tree using rpart.plot function in rpart.plot[6] package
library(rpart.plot)
prp(LTRCART.obj, roundint=FALSE)

## Plot the fitted LTRCIT tree
plot(LTRCIT.obj)
```

```{r}
library(partykit)
LTRCART.obj.party <- as.party(LTRCART.obj) 
LTRCART.obj.party$fitted[["(response)"]]<- Surv(Train$age, Train$End, Train$death)
plot(LTRCART.obj.party)
```

```{r}
## predict median survival time on test data using fitted LTRCIT tree
LTRCIT.pred <- predict(LTRCIT.obj, newdata=Test, type = "response")
head(LTRCIT.pred)
```

```{r}
## predict Kaplan Meier survival curve on test data
## return a list of survfit objects -- the predicted KM curves
LTRCIT.pred <- predict(LTRCIT.obj, newdata=Test, type = "prob")
head(LTRCIT.pred,2)
```

```{r}
## Predict relative risk on test set
LTRCART.pred <- predict(LTRCART.obj, newdata=Test)
head(LTRCART.pred)
```

```{r}
## Predict median survival time and Kaplan Meier survival curve
## on test data using Pred.rpart
LTRCART.pred <- Pred.rpart(Surv(age, End, death) ~ sex + FLC + creatinine, Train, Test)
head(LTRCART.pred$KMcurves, 2)  ## list of predicted KM curves
```

```{r}
head(LTRCART.pred$Medians)  ## vector of predicted median survival time
```

## Examples of fitting survival trees with time-varying covariates

### The Mayo Clinic Primary Biliary Cirrhosis Data example

```{r}
set.seed(0)
library(survival)
## Create the start-stop-event triplet needed for coxph and LTRC trees
first <- with(pbcseq, c(TRUE, diff(id) !=0)) #first id for each subject
last <- c(first[-1], TRUE) #last id
time1 <- with(pbcseq, ifelse(first, 0, day))
time2 <- with(pbcseq, ifelse(last, futime, c(day[-1], 0)))
event <- with(pbcseq, ifelse(last, status, 0))
event <- 1*(event==2)

pbcseq$time1 <- time1
pbcseq$time2 <- time2
pbcseq$event <-  event

head(pbcseq)
```

```{r}
## Fit the Cox model and LTRC trees with time-varying covariates
fit.cox <- coxph(Surv(time1, time2, event) ~ age + sex + log(bili), pbcseq)
LTRCIT.fit <- LTRCIT(Surv(time1, time2, event) ~ age + sex + log(bili), pbcseq)
LTRCART.fit <- LTRCART(Surv(time1, time2, event) ~ age + sex + log(bili), pbcseq)

## Result of the Cox model with time-varying covariates
fit.cox 
```

```{r}
## plots of fitted survival trees with time-varying covariates
prp(LTRCART.fit, type=0, roundint=FALSE)
plot(LTRCIT.fit)
```

### The Stanford Heart Transplant data example

```{r}
library(survival)
### transform the wide format data into the long format data using tmerge function
### from survival package on Stanford Heart Transplant data
jasa$subject <- 1:nrow(jasa)

tdata <- with(jasa, data.frame(subject = subject,
    futime= pmax(.5, fu.date - accept.dt),
    txtime= ifelse(tx.date== fu.date,
    (tx.date -accept.dt) -.5,
    (tx.date - accept.dt)),
    fustat = fustat))

sdata <- tmerge(jasa, tdata, id=subject,death = event(futime, fustat),trt = tdc(txtime), options= list(idname="subject"))

sdata$age <- sdata$age - 48

sdata$year <- as.numeric(sdata$accept.dt - as.Date("1967-10-01"))/365.25

head(sdata)
```

```{r}
Cox.fit <- coxph(Surv(tstart, tstop, death) ~ age + surgery, data= sdata)
LTRCART.fit <- LTRCART(Surv(tstart, tstop, death) ~ age + transplant, data = sdata)
LTRCIT.fit <- LTRCIT(Surv(tstart, tstop, death) ~ age + transplant, data = sdata)

## results
Cox.fit
```

```{r}
prp(LTRCART.fit, roundint=FALSE)
plot(LTRCIT.fit)
```

## SurvCART: Constructing Survival Tree in R

- Reference : https://www.r-bloggers.com/2022/01/survcart-constructing-survival-tree-in-r/

### The GBSG2 dataset

```{r}
library(LongCART)

data(GBSG2)

GBSG2$horTh1<- as.numeric(GBSG2$horTh)
GBSG2$tgrade1<- as.numeric(GBSG2$tgrade)
GBSG2$menostat1<- as.numeric(GBSG2$menostat)
GBSG2$subjid<- 1:nrow(GBSG2)

head(GBSG2)

```

```{r warning = FALSE, message = FALSE}
out.tree<- SurvCART(data=GBSG2, patid="subjid", 
                censorvar="cens", timevar="time", 
                gvars=c('horTh1', 'age', 'menostat1', 'tsize',          
                        'tgrade1', 'pnodes', 'progrec', 'estrec'),  
                tgvars=c(0,1,0,1,0,1,1,1), 
                time.dist="exponential", 
                event.ind=1, alpha=0.05, minsplit=80, minbucket=40, 
                print=TRUE)
```

```{r}
par(xpd = TRUE)
plot(out.tree, compress = TRUE)
text(out.tree, use.n = TRUE)
```

```{r}
KMPlot.SurvCART(out.tree, scale.time=365.25, type=1)
```

