---
title: "Time dependent ROC curve with Cox regression (ggplot2)"
author: "Hwang Seong-Yun"
date: '2022 4 1 '
output: html_document
---

# Time dependent ROC curve with Cox regression


## Loading packages

```{r}
library(survival)
library(timeROC)
library(timereg)
```


## pbc data

```{r}
data(pbc)
pbc<-pbc[,-1] # delete ID
pbc$status<-as.numeric(pbc$status==2) # create event indicator: 1 for death, 0 for censored
head(pbc)
```

### Calculate risk score with Cox proportional hazard model

```{r}
cox.pbc <- coxph(Surv(time,status) ~ ., data=pbc)
summary(cox.pbc)
risk.pbc <- predict(object=cox.pbc, newdata=pbc, type="risk") # risk score
pbc$risk <- risk.pbc
pbc <- pbc[!is.na(pbc$risk),]
```

### Time dependent AUC with risk score

```{r}
ROC.risk.marginal <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal

# plot all ROC curves 
par(mar=c(4,4,3,1))
plot(ROC.risk.marginal,time=1191,lwd=2,title=FALSE)
plot(ROC.risk.marginal,time=1786,col="blue",add=TRUE,lwd=2,lty=2)
plot(ROC.risk.marginal,time=2689,col="black",add=TRUE,lwd=2,lty=3)
# add legend
legend("bottomright",c("t=1191","t=1786","t=2689"),
       col=c("red","blue","black"),lty=1:3)
```


### time VS AUC

#### Make density of time 

```{r}
time.density <- density(ROC.risk.marginal[["times"]])
data1 <- data.frame(time=time.density[["x"]], density=time.density[["y"]])
data1$density.normal1 <- 0.9*(time.density[["y"]]-min(time.density[["y"]]))/(max(time.density[["y"]])-min(time.density[["y"]]))
data1$density.normal2 <- 0.9*(max(na.omit(ROC.risk.marginal[["AUC"]]))-min(na.omit(ROC.risk.marginal[["AUC"]])))*(time.density[["y"]]-min(time.density[["y"]]))/(max(time.density[["y"]])-min(time.density[["y"]]))+min(na.omit(ROC.risk.marginal[["AUC"]])) 
```

#### Plot of density of time 

```{r}
par(mar=c(4,4,3,1))
plot(time.density)
rug(jitter(ROC.risk.marginal[["times"]]))
```

#### Coloring status

```{r}
pbc$status.color <- ifelse(pbc$status==1,"seagreen","darkred")
pbc$status1 <- ifelse(pbc$status==1,1,min(na.omit(ROC.risk.marginal[["AUC"]])))
```

#### Plot of Time dependent AUC with Adjusted Density of time and Status of censoring

```{r}
par(mar=c(4,4,3,1))
plot(x=ROC.risk.marginal[["times"]],y=ROC.risk.marginal[["AUC"]],
     type="l",xlab="times",ylab="AUC",lwd=2,
     main="Time dependent AUC with risk score (pbc data)")
lines(x=data1$time,y=data1$density.normal2,col="steelblue",type="l",lwd=2)
points(x=pbc$time,y=pbc$status1,col=pbc$status.color,pch=20)
abline(v=seq(0,5000,500),col="lightgray",lty=1)
rug(jitter(ROC.risk.marginal[["times"]]))
legend(450,0.84,c("AUC","Density of time(adjusted)","censored(status=0)"),
       col=c("black","steelblue","darkred"),lwd=c(2,2,2),lty=c(1,1,2))
```

```{r}
par(mar=c(4,4,3,1))
plot(x=ROC.risk.marginal[["times"]],y=ROC.risk.marginal[["AUC"]],
     type="l",xlab="times",ylab="AUC",ylim=c(0,1),lwd=2,
     main="Time dependent AUC with risk score (pbc data) / ylim=c(0,1)")
lines(x=data1$time,y=data1$density.normal1,col="steelblue",type="l",lwd=2)
points(x=pbc$time,y=pbc$status,col=pbc$status.color,pch=20)
abline(v=seq(0,5000,500),col="lightgray",lty=1)
rug(jitter(ROC.risk.marginal[["times"]]))
legend(750,0.5,c("AUC","Density of time(adjusted)","censored(status=0)"),
       col=c("black","steelblue","darkred"),lwd=c(2,2,2),lty=c(1,1,2))
```

#### Using ggplot2

```{r}
library(ggplot2)
data2 <- data.frame(time=ROC.risk.marginal[["times"]],
                    AUC=ROC.risk.marginal[["AUC"]])
ggplot() +
  geom_line(data=data2, aes(x=time, y=AUC), size=1.5, colour='black') +
  geom_point(data=pbc, aes(x=time, y=status1, color=as.factor(status))) +
  geom_line(data=data1, aes(x=time, y=density.normal2), size=1, colour='blue') +
  scale_x_continuous(limits = c(0, max(pbc$time))) +
  labs(title="Time dependent AUC with risk score (pbc data)") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Counting censored data

```{r}
library(tidyverse)
pbc.censor <- pbc[pbc$status==0,]
pbc.censor <- pbc.censor %>%
  mutate(n.cen=cut(pbc.censor$time, 
                   breaks=seq(from=0,to=5000,by=500)))
table(pbc.censor$n.cen)
```


### Variable selection (backward)

#### 유의확률을 기준으로 유의하지 않은 변수들을 순차적으로 제거

#### All explanatory variables

```{r}
data(pbc)
pbc<-pbc[,-1] # delete ID
pbc$status<-as.numeric(pbc$status==2) # create event indicator: 1 for death, 0 for censored
cox.pbc0 <- coxph(Surv(time,status) ~ trt+age+sex+ascites+hepato+spiders+edema
                 +bili+chol+albumin+copper+alk.phos+ast+trig
                 +platelet+protime+stage, data=pbc)
summary(cox.pbc0)
risk.pbc0 <- predict(object=cox.pbc0, newdata=pbc, type="risk") # risk score
pbc$risk0 <- risk.pbc0
pbc <- pbc[!is.na(pbc$risk0),]
```

#### Except explanatory variables with (p-value > 0.5)

```{r}
cox.pbc1 <- coxph(Surv(time,status) ~ age+sex+edema
                 +bili+chol+albumin+copper+ast+trig
                 +platelet+protime+stage, data=pbc)
summary(cox.pbc1)
risk.pbc1 <- predict(object=cox.pbc1, newdata=pbc, type="risk") # risk score
pbc$risk1 <- risk.pbc1
```

#### Except explanatory variables with (p-value > 0.5)

```{r}
cox.pbc2 <- coxph(Surv(time,status) ~ age+sex+edema
                 +bili+chol+albumin+copper+ast+protime+stage, data=pbc)
summary(cox.pbc2)
risk.pbc2 <- predict(object=cox.pbc2, newdata=pbc, type="risk") # risk score
pbc$risk2 <- risk.pbc2
```

#### Except explanatory variables with (p-value > 0.3)

```{r}
cox.pbc3 <- coxph(Surv(time,status) ~ age+edema
                 +bili+chol+albumin+copper+ast+protime+stage, data=pbc)
summary(cox.pbc3)
risk.pbc3 <- predict(object=cox.pbc3, newdata=pbc, type="risk") # risk score
pbc$risk3 <- risk.pbc3
```

#### Except explanatory variables with (p-value > 0.1)

```{r}
cox.pbc4 <- coxph(Surv(time,status) ~ age+edema+bili+albumin+copper+ast+protime+stage, data=pbc)
summary(cox.pbc4)
risk.pbc4 <- predict(object=cox.pbc4, newdata=pbc, type="risk") # risk score
pbc$risk4 <- risk.pbc4
```

#### Time dependent AUC with Backward elimination

```{r}
ROC.risk.marginal0 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk0,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal0
ROC.risk.marginal1 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk1,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal1
ROC.risk.marginal2 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk2,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal2
ROC.risk.marginal3 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk3,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal3
ROC.risk.marginal4 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal4

data3 <- data.frame(time=ROC.risk.marginal0[["times"]],
                    cox.pbc0=ROC.risk.marginal0[["AUC"]],
                    cox.pbc1=ROC.risk.marginal1[["AUC"]],
                    cox.pbc2=ROC.risk.marginal2[["AUC"]],
                    cox.pbc3=ROC.risk.marginal3[["AUC"]],
                    cox.pbc4=ROC.risk.marginal4[["AUC"]])
library(reshape)
library(plyr)
data3.melt <- melt(data = data3, 
                   id.vars = "time", 
                   measure.vars = c("cox.pbc0","cox.pbc1","cox.pbc2",
                                    "cox.pbc3","cox.pbc4"))
data3.melt <- rename(data3.melt,
                     replace=c("variable"="model",
                               "value"="AUC"))
```

#### 유의한 설명변수들만 포함한 모형 cox.pbc4에서 각 변수를 하나씩 제거

#### Except each explanatory variables in model cox.pbc4

```{r}
#Except age
cox.pbc4_1 <- coxph(Surv(time,status) ~ edema+bili+albumin+copper+ast+protime+stage, data=pbc)
summary(cox.pbc4_1)
risk.pbc4_1 <- predict(object=cox.pbc4_1, newdata=pbc, type="risk") # risk score
pbc$risk4_1 <- risk.pbc4_1
ROC.risk.marginal4_1 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_1,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal4_1

#Except edema
cox.pbc4_2 <- coxph(Surv(time,status) ~ age+bili+albumin+copper+ast+protime+stage, data=pbc)
summary(cox.pbc4_2)
risk.pbc4_2 <- predict(object=cox.pbc4_2, newdata=pbc, type="risk") # risk score
pbc$risk4_2 <- risk.pbc4_2
ROC.risk.marginal4_2 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_2,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal4_2

#Except bili
cox.pbc4_3 <- coxph(Surv(time,status) ~ age+edema+albumin+copper+ast+protime+stage, data=pbc)
summary(cox.pbc4_3)
risk.pbc4_3 <- predict(object=cox.pbc4_3, newdata=pbc, type="risk") # risk score
pbc$risk4_3 <- risk.pbc4_3
ROC.risk.marginal4_3 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_3,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal4_3

#Except albumin
cox.pbc4_4 <- coxph(Surv(time,status) ~ age+edema+bili+copper+ast+protime+stage, data=pbc)
summary(cox.pbc4_4)
risk.pbc4_4 <- predict(object=cox.pbc4_4, newdata=pbc, type="risk") # risk score
pbc$risk4_4 <- risk.pbc4_4
ROC.risk.marginal4_4 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_4,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal4_4

#Except copper
cox.pbc4_5 <- coxph(Surv(time,status) ~ age+edema+bili+albumin+ast+protime+stage, data=pbc)
summary(cox.pbc4_5)
risk.pbc4_5 <- predict(object=cox.pbc4_5, newdata=pbc, type="risk") # risk score
pbc$risk4_5 <- risk.pbc4_5
ROC.risk.marginal4_5 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_5,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal4_5

#Except ast
cox.pbc4_6 <- coxph(Surv(time,status) ~ age+edema+bili+albumin+copper+protime+stage, data=pbc)
summary(cox.pbc4_6)
risk.pbc4_6 <- predict(object=cox.pbc4_6, newdata=pbc, type="risk") # risk score
pbc$risk4_6 <- risk.pbc4_6
ROC.risk.marginal4_6 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_6,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal4_6

#Except protime
cox.pbc4_7 <- coxph(Surv(time,status) ~ age+edema+bili+albumin+copper+ast+stage, data=pbc)
summary(cox.pbc4_7)
risk.pbc4_7 <- predict(object=cox.pbc4_7, newdata=pbc, type="risk") # risk score
pbc$risk4_7 <- risk.pbc4_7
ROC.risk.marginal4_7 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_7,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal4_7

#Except stage
cox.pbc4_8 <- coxph(Surv(time,status) ~ age+edema+bili+albumin+copper+ast+protime, data=pbc)
summary(cox.pbc4_8)
risk.pbc4_8 <- predict(object=cox.pbc4_8, newdata=pbc, type="risk") # risk score
pbc$risk4_8 <- risk.pbc4_8
ROC.risk.marginal4_8 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_8,
                         cause=1,weighting="marginal",
                         times=pbc$time)
ROC.risk.marginal4_8

data4 <- data.frame(time=ROC.risk.marginal4[["times"]],
                    cox.pbc4=ROC.risk.marginal4[["AUC"]],
                    cox.pbc4_1=ROC.risk.marginal4_1[["AUC"]],
                    cox.pbc4_2=ROC.risk.marginal4_2[["AUC"]],
                    cox.pbc4_3=ROC.risk.marginal4_3[["AUC"]],
                    cox.pbc4_4=ROC.risk.marginal4_4[["AUC"]],
                    cox.pbc4_5=ROC.risk.marginal4_5[["AUC"]],
                    cox.pbc4_6=ROC.risk.marginal4_6[["AUC"]],
                    cox.pbc4_7=ROC.risk.marginal4_7[["AUC"]],
                    cox.pbc4_8=ROC.risk.marginal4_8[["AUC"]])
data4.melt <- melt(data = data4, 
                   id.vars = "time", 
                   measure.vars = c("cox.pbc4","cox.pbc4_1","cox.pbc4_2",
                                    "cox.pbc4_3","cox.pbc4_4","cox.pbc4_5",
                                    "cox.pbc4_6","cox.pbc4_7","cox.pbc4_8"))
data4.melt <- rename(data4.melt,
                     replace=c("variable"="model",
                               "value"="AUC"))
```

```{r}
time.density <- density(ROC.risk.marginal[["times"]])
data1 <- data.frame(time=time.density[["x"]], density=time.density[["y"]])
data1$density.normal1 <- (time.density[["y"]]-min(time.density[["y"]]))/(max(time.density[["y"]])-min(time.density[["y"]]))
data1$density.normal2 <- (max(na.omit(ROC.risk.marginal0[["AUC"]]))-min(na.omit(ROC.risk.marginal0[["AUC"]])))*(time.density[["y"]]-min(time.density[["y"]]))/(max(time.density[["y"]])-min(time.density[["y"]]))+min(na.omit(ROC.risk.marginal0[["AUC"]]))
```

```{r}
pbc$status.color <- ifelse(pbc$status==1,"seagreen","darkred")
pbc$status1 <- ifelse(pbc$status==1,1,0.7)
```

#### Plot of Time dependent AUC with Adjusted Density of time and Status of censoring (cox.pbc0~4)

```{r}
par(mar=c(4,4,3,1))
plot(x=ROC.risk.marginal0[["times"]],y=ROC.risk.marginal0[["AUC"]],
     type="l",xlab="times",ylab="AUC",lwd=1,lty=1,col="black",ylim=c(0.7,1),
     main="Time dependent AUC with risk score (pbc data)")
lines(x=ROC.risk.marginal1[["times"]],y=ROC.risk.marginal1[["AUC"]],
     type="l",lwd=1,lty=1,col="red")
lines(x=ROC.risk.marginal2[["times"]],y=ROC.risk.marginal2[["AUC"]],
     type="l",lwd=1,lty=1,col="blue")
lines(x=ROC.risk.marginal3[["times"]],y=ROC.risk.marginal3[["AUC"]],
     type="l",lwd=1,lty=1,col="green")
lines(x=ROC.risk.marginal4[["times"]],y=ROC.risk.marginal4[["AUC"]],
     type="l",lwd=1,lty=1,col="yellow")
lines(x=data1$time,y=data1$density.normal2,col="steelblue",type="l",lwd=2)
points(x=pbc$time,y=pbc$status1,col=pbc$status.color,pch=20)
abline(v=seq(0,5000,500),col="lightgray",lty=1)
rug(jitter(ROC.risk.marginal[["times"]]))
legend(400,0.85,c("AUC(cox.pbc0)","AUC(cox.pbc1)","AUC(cox.pbc2)","AUC(cox.pbc3)","AUC(cox.pbc4)","Density of time(adjusted)","censored(status=0)"),col=c("black","red","blue","green","yellow","steelblue","darkred"),lwd=c(1,1,1,1,1,1,1),lty=c(1,1,1,1,1,1,2))
```

#### Plot without legend

```{r}
par(mar=c(4,4,3,1))
plot(x=ROC.risk.marginal0[["times"]],y=ROC.risk.marginal0[["AUC"]],
     type="l",xlab="times",ylab="AUC",lwd=1,lty=1,col="black",ylim=c(0.7,1),
     main="Time dependent AUC with risk score (pbc data)")
lines(x=ROC.risk.marginal1[["times"]],y=ROC.risk.marginal1[["AUC"]],
     type="l",lwd=1,lty=1,col="red")
lines(x=ROC.risk.marginal2[["times"]],y=ROC.risk.marginal2[["AUC"]],
     type="l",lwd=1,lty=1,col="blue")
lines(x=ROC.risk.marginal3[["times"]],y=ROC.risk.marginal3[["AUC"]],
     type="l",lwd=1,lty=1,col="green")
lines(x=ROC.risk.marginal4[["times"]],y=ROC.risk.marginal4[["AUC"]],
     type="l",lwd=1,lty=1,col="yellow")
lines(x=data1$time,y=data1$density.normal2,col="steelblue",type="l",lwd=2)
points(x=pbc$time,y=pbc$status1,col=pbc$status.color,pch=20)
abline(v=seq(0,5000,500),col="lightgray",lty=1)
rug(jitter(ROC.risk.marginal[["times"]]))
```

#### Using ggplot2 (cox.pbc0~4)

```{r}
pbc$status1 <- ifelse(pbc$status==1,1,0.7)
ggplot() +
  geom_line(data=data3.melt, aes(x=time, y=AUC,
                            group=model, color=model)) +
  geom_point(data=pbc, aes(x=time, y=status1, color=as.factor(status))) +
  geom_line(data=data1, aes(x=time, y=density.normal2), size=1, colour='black') +
  scale_x_continuous(limits = c(0, max(pbc$time))) +
  labs(title="Time dependent AUC with risk score (cox.pbc0~4)",
       xlab="time", ylab="AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.pbc0~4)

```{r}
mean(ROC.risk.marginal0[["AUC"]],na.rm=TRUE)
mean(ROC.risk.marginal1[["AUC"]],na.rm=TRUE)
mean(ROC.risk.marginal2[["AUC"]],na.rm=TRUE)
mean(ROC.risk.marginal3[["AUC"]],na.rm=TRUE)
mean(ROC.risk.marginal4[["AUC"]],na.rm=TRUE)
```

#### Plot of Time dependent AUC with Adjusted Density of time and Status of censoring (cox.pbc4~4_8)

```{r}
par(mar=c(4,4,3,1))
plot(x=ROC.risk.marginal4[["times"]],y=ROC.risk.marginal4[["AUC"]],
     type="l",xlab="times",ylab="AUC",lwd=3,lty=1,col="black",ylim=c(0.7,1),
     main="Time dependent AUC with risk score (pbc data)")
lines(x=ROC.risk.marginal4_1[["times"]],y=ROC.risk.marginal4_1[["AUC"]],
     type="l",lwd=1,lty=1,col="red") #Except age
lines(x=ROC.risk.marginal4_2[["times"]],y=ROC.risk.marginal4_2[["AUC"]],
     type="l",lwd=1,lty=1,col="blue") #Except edema
lines(x=ROC.risk.marginal4_3[["times"]],y=ROC.risk.marginal4_3[["AUC"]],
     type="l",lwd=1,lty=1,col="green") #Except bili
lines(x=ROC.risk.marginal4_4[["times"]],y=ROC.risk.marginal4_4[["AUC"]],
     type="l",lwd=1,lty=1,col="yellow") #Except albumin
lines(x=ROC.risk.marginal4_5[["times"]],y=ROC.risk.marginal4_5[["AUC"]],
     type="l",lwd=1,lty=1,col="brown") #Except copper  
lines(x=ROC.risk.marginal4_6[["times"]],y=ROC.risk.marginal4_6[["AUC"]],
     type="l",lwd=1,lty=1,col="purple") #Except ast
lines(x=ROC.risk.marginal4_7[["times"]],y=ROC.risk.marginal4_7[["AUC"]],
     type="l",lwd=1,lty=1,col="navy") #Except protime
lines(x=ROC.risk.marginal4_8[["times"]],y=ROC.risk.marginal4_8[["AUC"]],
     type="l",lwd=1,lty=1,col="orange")#Except stage
lines(x=data1$time,y=data1$density.normal2,col="steelblue",type="l",lwd=2)
points(x=pbc$time,y=pbc$status1,col=pbc$status.color,pch=20)
abline(v=seq(0,5000,500),col="lightgray",lty=1)
rug(jitter(ROC.risk.marginal[["times"]]))
legend("topright",c("AUC(cox.pbc4)","AUC(cox.pbc4_1)","AUC(cox.pbc4_2)","AUC(cox.pbc4_3)","AUC(cox.pbc4_4)","AUC(cox.pbc4_5)","AUC(cox.pbc4_6)","AUC(cox.pbc4_7)","AUC(cox.pbc4_8)","Density of time(adjusted)","censored(status=0)"),col=c("black","red","blue","green","yellow","brown","purple","navy","orange","steelblue","darkred"),lwd=c(1,1,1,1,1,1,1,1,1,1,1),lty=c(1,1,1,1,1,1,1,1,1,1,2))
```

#### Plot without legend

```{r}
par(mar=c(4,4,3,1))
plot(x=ROC.risk.marginal4[["times"]],y=ROC.risk.marginal4[["AUC"]],
     type="l",xlab="times",ylab="AUC",lwd=3,lty=1,col="black",ylim=c(0.7,1),
     main="Time dependent AUC with risk score (pbc data)")
lines(x=ROC.risk.marginal4_1[["times"]],y=ROC.risk.marginal4_1[["AUC"]],
     type="l",lwd=1,lty=1,col="red") #Except age
lines(x=ROC.risk.marginal4_2[["times"]],y=ROC.risk.marginal4_2[["AUC"]],
     type="l",lwd=1,lty=1,col="blue") #Except edema
lines(x=ROC.risk.marginal4_3[["times"]],y=ROC.risk.marginal4_3[["AUC"]],
     type="l",lwd=1,lty=1,col="green") #Except bili
lines(x=ROC.risk.marginal4_4[["times"]],y=ROC.risk.marginal4_4[["AUC"]],
     type="l",lwd=1,lty=1,col="yellow") #Except albumin
lines(x=ROC.risk.marginal4_5[["times"]],y=ROC.risk.marginal4_5[["AUC"]],
     type="l",lwd=1,lty=1,col="brown") #Except copper  
lines(x=ROC.risk.marginal4_6[["times"]],y=ROC.risk.marginal4_6[["AUC"]],
     type="l",lwd=1,lty=1,col="purple") #Except ast
lines(x=ROC.risk.marginal4_7[["times"]],y=ROC.risk.marginal4_7[["AUC"]],
     type="l",lwd=1,lty=1,col="navy") #Except protime
lines(x=ROC.risk.marginal4_8[["times"]],y=ROC.risk.marginal4_8[["AUC"]],
     type="l",lwd=1,lty=1,col="orange")#Except stage
lines(x=data1$time,y=data1$density.normal2,col="steelblue",type="l",lwd=2)
points(x=pbc$time,y=pbc$status1,col=pbc$status.color,pch=20)
abline(v=seq(0,5000,500),col="lightgray",lty=1)
rug(jitter(ROC.risk.marginal[["times"]]))
```


#### Using ggplot2 (cox.pbc4~4_8)

```{r}
pbc$status1 <- ifelse(pbc$status==1,1,0.65)
ggplot() +
  geom_line(data=data4.melt, aes(x=time, y=AUC,
                            group=model, color=model)) +
  geom_point(data=pbc, aes(x=time, y=status1, color=as.factor(status))) +
  geom_line(data=data1, aes(x=time, y=density.normal2), size=1, colour='black') +
  scale_x_continuous(limits = c(0, max(pbc$time))) +
  labs(title="Time dependent AUC with risk score (cox.pbc4~4_8)",
       xlab="time", ylab="AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.pbc4~4_8)

```{r}
mean(ROC.risk.marginal4[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.marginal4_1[["AUC"]],na.rm=TRUE) #Except age
mean(ROC.risk.marginal4_2[["AUC"]],na.rm=TRUE) #Except edema
mean(ROC.risk.marginal4_3[["AUC"]],na.rm=TRUE) #Except bili
mean(ROC.risk.marginal4_4[["AUC"]],na.rm=TRUE) #Except albumin 
mean(ROC.risk.marginal4_5[["AUC"]],na.rm=TRUE) #Except copper 
mean(ROC.risk.marginal4_6[["AUC"]],na.rm=TRUE) #Except ast 
mean(ROC.risk.marginal4_7[["AUC"]],na.rm=TRUE) #Except protime 
mean(ROC.risk.marginal4_8[["AUC"]],na.rm=TRUE) #Except stage 
```


### Backward Elimination ver2

#### 유의한 설명변수들만 포함한 모형 cox.pbc4에서 유의확률을 기준으로 가장 유의한 변수를 순차적으로 하나씩 제거

#### All significant explanatory variables 

```{r}
cox.pbc4 <- coxph(Surv(time,status) ~ age+edema+bili+albumin+copper+ast+protime+stage, data=pbc)
summary(cox.pbc4)
risk.pbc4 <- predict(object=cox.pbc4, newdata=pbc, type="risk") # risk score
pbc$risk4 <- risk.pbc4
ROC.risk.marginal4 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4,
                         cause=1,weighting="marginal",
                         times=pbc$time)
```

#### Except bili

```{r}
cox.pbc4_11 <- coxph(Surv(time,status) ~ age+edema+albumin+copper+ast+protime+stage, data=pbc)
summary(cox.pbc4_11)
risk.pbc4_11 <- predict(object=cox.pbc4_11, newdata=pbc, type="risk") # risk score
pbc$risk4_11 <- risk.pbc4_11
ROC.risk.marginal4_11 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_11,
                         cause=1,weighting="marginal",
                         times=pbc$time)
```

#### Except bili, ast

```{r}
cox.pbc4_12 <- coxph(Surv(time,status) ~ age+edema+albumin+copper+protime+stage, data=pbc)
summary(cox.pbc4_12)
risk.pbc4_12 <- predict(object=cox.pbc4_12, newdata=pbc, type="risk") # risk score
pbc$risk4_12 <- risk.pbc4_12
ROC.risk.marginal4_12 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_12,
                         cause=1,weighting="marginal",
                         times=pbc$time)
```

#### Except bili, ast, copper

```{r}
cox.pbc4_13 <- coxph(Surv(time,status) ~ age+edema+albumin+protime+stage, data=pbc)
summary(cox.pbc4_13)
risk.pbc4_13 <- predict(object=cox.pbc4_13, newdata=pbc, type="risk") # risk score
pbc$risk4_13 <- risk.pbc4_13
ROC.risk.marginal4_13 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_13,
                         cause=1,weighting="marginal",
                         times=pbc$time)
```

#### Except bili, ast, copper, edema

```{r}
cox.pbc4_14 <- coxph(Surv(time,status) ~ age+albumin+protime+stage, data=pbc)
summary(cox.pbc4_14)
risk.pbc4_14 <- predict(object=cox.pbc4_14, newdata=pbc, type="risk") # risk score
pbc$risk4_14 <- risk.pbc4_14
ROC.risk.marginal4_14 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_14,
                         cause=1,weighting="marginal",
                         times=pbc$time)
```

#### Except bili, ast, copper, edema, protime

```{r}
cox.pbc4_15 <- coxph(Surv(time,status) ~ age+albumin+stage, data=pbc)
summary(cox.pbc4_15)
risk.pbc4_15 <- predict(object=cox.pbc4_15, newdata=pbc, type="risk") # risk score
pbc$risk4_15 <- risk.pbc4_15
ROC.risk.marginal4_15 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_15,
                         cause=1,weighting="marginal",
                         times=pbc$time)
```

#### Except bili, ast, copper, edema, protime, albumin

```{r}
cox.pbc4_16 <- coxph(Surv(time,status) ~ age+stage, data=pbc)
summary(cox.pbc4_16)
risk.pbc4_16 <- predict(object=cox.pbc4_16, newdata=pbc, type="risk") # risk score
pbc$risk4_16 <- risk.pbc4_16
ROC.risk.marginal4_16 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_16,
                         cause=1,weighting="marginal",
                         times=pbc$time)
```

#### Except bili, ast, copper, edema, protime, albumin, stage

```{r}
cox.pbc4_17 <- coxph(Surv(time,status) ~ age, data=pbc)
summary(cox.pbc4_17)
risk.pbc4_17 <- predict(object=cox.pbc4_17, newdata=pbc, type="risk") # risk score
pbc$risk4_17 <- risk.pbc4_17
ROC.risk.marginal4_17 <- timeROC(T=pbc$time,
                         delta=pbc$status,marker=pbc$risk4_17,
                         cause=1,weighting="marginal",
                         times=pbc$time)
```

#### Make AUC data

```{r}
data4_1 <- data.frame(time=ROC.risk.marginal4[["times"]],
                    cox.pbc4=ROC.risk.marginal4[["AUC"]],
                    cox.pbc4_11=ROC.risk.marginal4_11[["AUC"]],
                    cox.pbc4_12=ROC.risk.marginal4_12[["AUC"]],
                    cox.pbc4_13=ROC.risk.marginal4_13[["AUC"]],
                    cox.pbc4_14=ROC.risk.marginal4_14[["AUC"]],
                    cox.pbc4_15=ROC.risk.marginal4_15[["AUC"]],
                    cox.pbc4_16=ROC.risk.marginal4_16[["AUC"]],
                    cox.pbc4_17=ROC.risk.marginal4_17[["AUC"]])
data4_1.melt <- melt(data = data4_1, 
                   id.vars = "time", 
                   measure.vars = c("cox.pbc4","cox.pbc4_11","cox.pbc4_12",
                                    "cox.pbc4_13","cox.pbc4_14","cox.pbc4_15",
                                    "cox.pbc4_16","cox.pbc4_17"))
data4_1.melt <- rename(data4_1.melt,
                     replace=c("variable"="model",
                               "value"="AUC"))

time.density <- density(ROC.risk.marginal[["times"]])
data1 <- data.frame(time=time.density[["x"]], density=time.density[["y"]])
data1$density.normal1 <- (time.density[["y"]]-min(time.density[["y"]]))/(max(time.density[["y"]])-min(time.density[["y"]]))
data1$density.normal2 <- (max(na.omit(ROC.risk.marginal4_17[["AUC"]]))-min(na.omit(ROC.risk.marginal4_17[["AUC"]])))*(time.density[["y"]]-min(time.density[["y"]]))/(max(time.density[["y"]])-min(time.density[["y"]]))+min(na.omit(ROC.risk.marginal4_17[["AUC"]]))

pbc$status.color <- ifelse(pbc$status==1,"seagreen","darkred")
pbc$status1 <- ifelse(pbc$status==1,1,0.3)
```

#### Plot of Time dependent AUC with Adjusted Density of time and Status of censoring (cox.pbc4~4_17)

```{r}
par(mar=c(4,4,3,1))
plot(x=ROC.risk.marginal4[["times"]],y=ROC.risk.marginal4[["AUC"]],
     type="l",xlab="times",ylab="AUC",lwd=3,lty=1,col="black",ylim=c(0.3,1),
     main="Time dependent AUC with risk score (pbc data)")
lines(x=ROC.risk.marginal4_11[["times"]],y=ROC.risk.marginal4_11[["AUC"]],
     type="l",lwd=1,lty=1,col="red") #Except bili
lines(x=ROC.risk.marginal4_12[["times"]],y=ROC.risk.marginal4_12[["AUC"]],
     type="l",lwd=1,lty=1,col="blue") #Except bili, ast
lines(x=ROC.risk.marginal4_13[["times"]],y=ROC.risk.marginal4_13[["AUC"]],
     type="l",lwd=1,lty=1,col="green") #Except bili, ast, copper
lines(x=ROC.risk.marginal4_14[["times"]],y=ROC.risk.marginal4_14[["AUC"]],
     type="l",lwd=1,lty=1,col="yellow") #Except bili, ast, copper, edema
lines(x=ROC.risk.marginal4_15[["times"]],y=ROC.risk.marginal4_15[["AUC"]],
     type="l",lwd=1,lty=1,col="brown") #Except bili, ast, copper, edema, protime  
lines(x=ROC.risk.marginal4_16[["times"]],y=ROC.risk.marginal4_16[["AUC"]],
     type="l",lwd=1,lty=1,col="purple") #Except bili, ast, copper, edema, protime, albumin
lines(x=ROC.risk.marginal4_17[["times"]],y=ROC.risk.marginal4_17[["AUC"]],
     type="l",lwd=1,lty=1,col="navy") #Except bili, ast, copper, edema, protime, albumin, stage
lines(x=data1$time,y=data1$density.normal2,col="steelblue",type="l",lwd=2)
points(x=pbc$time,y=pbc$status1,col=pbc$status.color,pch=20)
abline(v=seq(0,5000,500),col="lightgray",lty=1)
rug(jitter(ROC.risk.marginal[["times"]]))
legend("topright",c("AUC(cox.pbc4)","AUC(cox.pbc4_11)","AUC(cox.pbc4_12)","AUC(cox.pbc4_13)","AUC(cox.pbc4_14)","AUC(cox.pbc4_15)","AUC(cox.pbc4_16)","AUC(cox.pbc4_17)","Density of time(adjusted)","censored(status=0)"),col=c("black","red","blue","green","yellow","brown","purple","navy","steelblue","darkred"),lwd=c(1,1,1,1,1,1,1,1,1,1),lty=c(1,1,1,1,1,1,1,1,1,2))
```

#### Plot without legend

```{r}
par(mar=c(4,4,3,1))
plot(x=ROC.risk.marginal4[["times"]],y=ROC.risk.marginal4[["AUC"]],
     type="l",xlab="times",ylab="AUC",lwd=3,lty=1,col="black",ylim=c(0.3,1),
     main="Time dependent AUC with risk score (pbc data)")
lines(x=ROC.risk.marginal4_11[["times"]],y=ROC.risk.marginal4_11[["AUC"]],
     type="l",lwd=1,lty=1,col="red") #Except bili
lines(x=ROC.risk.marginal4_12[["times"]],y=ROC.risk.marginal4_12[["AUC"]],
     type="l",lwd=1,lty=1,col="blue") #Except bili, ast
lines(x=ROC.risk.marginal4_13[["times"]],y=ROC.risk.marginal4_13[["AUC"]],
     type="l",lwd=1,lty=1,col="green") #Except bili, ast, copper
lines(x=ROC.risk.marginal4_14[["times"]],y=ROC.risk.marginal4_14[["AUC"]],
     type="l",lwd=1,lty=1,col="yellow") #Except bili, ast, copper, edema
lines(x=ROC.risk.marginal4_15[["times"]],y=ROC.risk.marginal4_15[["AUC"]],
     type="l",lwd=1,lty=1,col="brown") #Except bili, ast, copper, edema, protime  
lines(x=ROC.risk.marginal4_16[["times"]],y=ROC.risk.marginal4_16[["AUC"]],
     type="l",lwd=1,lty=1,col="purple") #Except bili, ast, copper, edema, protime, albumin
lines(x=ROC.risk.marginal4_17[["times"]],y=ROC.risk.marginal4_17[["AUC"]],
     type="l",lwd=1,lty=1,col="navy") #Except bili, ast, copper, edema, protime, albumin, stage
lines(x=data1$time,y=data1$density.normal2,col="steelblue",type="l",lwd=2)
points(x=pbc$time,y=pbc$status1,col=pbc$status.color,pch=20)
abline(v=seq(0,5000,500),col="lightgray",lty=1)
rug(jitter(ROC.risk.marginal[["times"]]))
```

#### Using ggplot2 (cox.pbc4~4_17)

```{r}
pbc$status1 <- ifelse(pbc$status==1,1,0.3)
ggplot() +
  geom_line(data=data4_1.melt, aes(x=time, y=AUC,
                            group=model, color=model)) +
  geom_point(data=pbc, aes(x=time, y=status1, color=as.factor(status))) +
  geom_line(data=data1, aes(x=time, y=density.normal2), size=1, colour='black') +
  scale_x_continuous(limits = c(0, max(pbc$time))) +
  labs(title="Time dependent AUC with risk score (cox.pbc4~4_17)",
       xlab="time", ylab="AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.pbc4~4_17)

```{r}
mean(ROC.risk.marginal4[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.marginal4_11[["AUC"]],na.rm=TRUE) #Except bili
mean(ROC.risk.marginal4_12[["AUC"]],na.rm=TRUE) #Except bili, ast
mean(ROC.risk.marginal4_13[["AUC"]],na.rm=TRUE) #Except bili, ast, copper
mean(ROC.risk.marginal4_14[["AUC"]],na.rm=TRUE) #Except bili, ast, copper, edema 
mean(ROC.risk.marginal4_15[["AUC"]],na.rm=TRUE) #Except bili, ast, copper, edema, protime 
mean(ROC.risk.marginal4_16[["AUC"]],na.rm=TRUE) #Except bili, ast, copper, edema, protime, albumin 
mean(ROC.risk.marginal4_17[["AUC"]],na.rm=TRUE) #Except bili, ast, copper, edema, protime, albumin, stage 
```





## Train vs Test

```{r}
data(pbc)
pbc<-pbc[,-1] # delete ID
pbc$status<-as.numeric(pbc$status==2) # create event indicator: 1 for death, 0 for censored
pbc <- na.omit(pbc)
```

### train:test = 7:3

### 임의로 pbc data를 훈련자료와 검증자료로 나누어서 분석

```{r}
set.seed(55364)
idx_train <- sample(1:nrow(pbc), round(0.7*nrow(pbc)))
pbc.train <- pbc[idx_train,]
pbc.test <- pbc[setdiff(x = 1:nrow(pbc), y = idx_train), ]
```

### Cox PH model with train data

#### 유의확률을 기준으로 유의하지 않은 변수들을 순차적으로 제거

#### Null model

```{r}
cox.train0 <- coxph(Surv(time,status) ~ trt+age+sex+ascites+hepato+spiders+edema
                 +bili+chol+albumin+copper+alk.phos+ast+trig
                 +platelet+protime+stage, data=pbc.train)
summary(cox.train0)
risk.train0 <- predict(object=cox.train0, newdata=pbc.test, type="risk") # risk score
pbc.test$risk0 <- risk.train0
ROC.risk.train0 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk0,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
ROC.risk.train0
```

#### Except explanatory variables with p-value > 0.5

```{r}
cox.train1 <- coxph(Surv(time,status) ~ age+sex+edema
                 +bili+chol+albumin+copper+ast
                 +protime+stage, data=pbc.train)
summary(cox.train1)
risk.train1 <- predict(object=cox.train1, newdata=pbc.test, type="risk") # risk score
pbc.test$risk1 <- risk.train1
ROC.risk.train1 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk1,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
ROC.risk.train1
```

#### Except explanatory variables with p-value > 0.2

```{r}
cox.train2 <- coxph(Surv(time,status) ~ age+bili+albumin+copper
                 +protime+stage, data=pbc.train)
summary(cox.train2)
risk.train2 <- predict(object=cox.train2, newdata=pbc.test, type="risk") # risk score
pbc.test$risk2 <- risk.train2
ROC.risk.train2 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk2,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
ROC.risk.train2
```

#### Except explanatory variables with p-value > 0.2

```{r}
cox.train3 <- coxph(Surv(time,status) ~ bili+albumin+copper
                 +protime+stage, data=pbc.train)
summary(cox.train3)
risk.train3 <- predict(object=cox.train3, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3 <- risk.train3
ROC.risk.train3 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
ROC.risk.train3
```

#### Make AUC data

```{r}
time.density <- density(ROC.risk.train0[["times"]])
data1 <- data.frame(time=time.density[["x"]], density=time.density[["y"]])
data1$density.adjusted <- (max(na.omit(ROC.risk.train0[["AUC"]]))-min(na.omit(ROC.risk.train0[["AUC"]])))*(time.density[["y"]]-min(time.density[["y"]]))/(max(time.density[["y"]])-min(time.density[["y"]]))+min(na.omit(ROC.risk.train0[["AUC"]]))
dt1 <- data.frame(time=ROC.risk.train0[["times"]],
                    cox.train0=ROC.risk.train0[["AUC"]],
                    cox.train1=ROC.risk.train1[["AUC"]],
                    cox.train2=ROC.risk.train2[["AUC"]],
                    cox.train3=ROC.risk.train3[["AUC"]])
dt1.melt <- melt(data = dt1, 
                   id.vars = "time", 
                   measure.vars = c("cox.train0","cox.train1","cox.train2",
                                    "cox.train3"))
dt1.melt <- rename(dt1.melt,
                     replace=c("variable"="model",
                               "value"="AUC"))
```

#### Time dependent AUC plot with ggplot2 (cox.train0~3)

```{r}
pbc.test$status1 <- ifelse(pbc.test$status==1,1,min(na.omit(ROC.risk.train0[["AUC"]])))
ggplot() +
  geom_line(data=dt1.melt, aes(x=time, y=AUC,
                            group=model, color=model)) +
  geom_point(data=pbc.test, aes(x=time, y=status1, color=as.factor(status))) +
  geom_line(data=data1, aes(x=time, y=density.adjusted), size=1, colour='black') +
  scale_x_continuous(limits = c(0, max(pbc.test$time))) +
  labs(title="Time dependent AUC with risk score (cox.train0~3)",
       xlab="time", ylab="AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.train0~3)

```{r}
mean(ROC.risk.train0[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.train1[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.train2[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.train3[["AUC"]],na.rm=TRUE) 
```

#### Counting censored data

```{r}
pbc.test.censor <- pbc.test[pbc.test$status==0,]
pbc.test.censor <- pbc.test.censor %>%
  mutate(n.cen=cut(pbc.test.censor$time, 
                   breaks=seq(from=0,to=5000,by=500)))
table(pbc.test.censor$n.cen)
```

#### 유의한 설명변수들만 포함한 모형 cox.train3에서 각 변수를 하나씩 제거

#### Except each explanatory variables in model cox.train3

```{r}
# Except bili
cox.train3_1 <- coxph(Surv(time,status) ~ albumin+copper
                 +protime+stage, data=pbc.train)
summary(cox.train3_1)
risk.train3_1 <- predict(object=cox.train3_1, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3_1 <- risk.train3_1
ROC.risk.train3_1 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3_1,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
ROC.risk.train3_1

# Except albumin
cox.train3_2 <- coxph(Surv(time,status) ~ bili+copper
                 +protime+stage, data=pbc.train)
summary(cox.train3_2)
risk.train3_2 <- predict(object=cox.train3_2, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3_2 <- risk.train3_2
ROC.risk.train3_2 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3_2,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
ROC.risk.train3_2

# Except copper
cox.train3_3 <- coxph(Surv(time,status) ~ bili+albumin
                 +protime+stage, data=pbc.train)
summary(cox.train3_3)
risk.train3_3 <- predict(object=cox.train3_3, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3_3 <- risk.train3_3
ROC.risk.train3_3 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3_3,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
ROC.risk.train3_3

# Except protime
cox.train3_4 <- coxph(Surv(time,status) ~ bili+albumin+copper
                 +stage, data=pbc.train)
summary(cox.train3_4)
risk.train3_4 <- predict(object=cox.train3_4, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3_4 <- risk.train3_4
ROC.risk.train3_4 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3_4,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
ROC.risk.train3_4

# Except stage
cox.train3_5 <- coxph(Surv(time,status) ~ bili+albumin+copper
                 +protime, data=pbc.train)
summary(cox.train3_5)
risk.train3_5 <- predict(object=cox.train3_5, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3_5 <- risk.train3_5
ROC.risk.train3_5 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3_5,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
ROC.risk.train3_5
```

#### Make AUC data

```{r}
time.density <- density(ROC.risk.train3[["times"]])
data1 <- data.frame(time=time.density[["x"]], density=time.density[["y"]])
data1$density.adjusted <- (max(na.omit(ROC.risk.train3[["AUC"]]))-min(na.omit(ROC.risk.train3[["AUC"]])))*(time.density[["y"]]-min(time.density[["y"]]))/(max(time.density[["y"]])-min(time.density[["y"]]))+min(na.omit(ROC.risk.train3[["AUC"]]))
dt2 <- data.frame(time=ROC.risk.train3[["times"]],
                    cox.train3=ROC.risk.train3[["AUC"]],
                    cox.train3_1=ROC.risk.train3_1[["AUC"]],
                    cox.train3_2=ROC.risk.train3_2[["AUC"]],
                    cox.train3_3=ROC.risk.train3_3[["AUC"]],
                    cox.train3_4=ROC.risk.train3_4[["AUC"]],
                    cox.train3_5=ROC.risk.train3_5[["AUC"]])
dt2.melt <- melt(data = dt2, 
                   id.vars = "time", 
                   measure.vars = c("cox.train3","cox.train3_1","cox.train3_2",
                                    "cox.train3_3","cox.train3_4","cox.train3_5"))
dt2.melt <- rename(dt2.melt,
                     replace=c("variable"="model",
                               "value"="AUC"))
```

#### Time dependent AUC plot with ggplot2 (cox.train3~3_5)

```{r}
pbc.test$status1 <- ifelse(pbc.test$status==1,1,min(na.omit(ROC.risk.train3[["AUC"]])))
ggplot() +
  geom_line(data=dt2.melt, aes(x=time, y=AUC,
                            group=model, color=model)) +
  geom_point(data=pbc.test, aes(x=time, y=status1, color=as.factor(status))) +
  geom_line(data=data1, aes(x=time, y=density.adjusted), size=1, colour='black') +
  scale_x_continuous(limits = c(0, max(pbc.test$time))) +
  labs(title="Time dependent AUC with risk score (cox.train3~3_5)",
       xlab="time", ylab="AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.train3~3_5)

```{r}
mean(ROC.risk.train3[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.train3_1[["AUC"]],na.rm=TRUE) # Except bili
mean(ROC.risk.train3_2[["AUC"]],na.rm=TRUE) # Except albumin
mean(ROC.risk.train3_3[["AUC"]],na.rm=TRUE) # Except copper
mean(ROC.risk.train3_4[["AUC"]],na.rm=TRUE) # Except protime 
mean(ROC.risk.train3_5[["AUC"]],na.rm=TRUE) # Except stage 
```

### Backward Elimination ver2

#### 유의한 설명변수들만 포함한 모형 cox.train3에서 유의확률을 기준으로 가장 유의한 변수를 순차적으로 하나씩 제거

#### All significant explanatory variables 

```{r}
cox.train3 <- coxph(Surv(time,status) ~ bili+albumin+copper+protime+stage, data=pbc.train)
summary(cox.train3)
risk.train3 <- predict(object=cox.train3, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3 <- risk.train3
ROC.risk.train3 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
```

#### Except bili

```{r}
cox.train3_11 <- coxph(Surv(time,status) ~ albumin+copper+protime+stage, data=pbc.train)
summary(cox.train3_11)
risk.train3_11 <- predict(object=cox.train3_11, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3_11 <- risk.train3_11
ROC.risk.train3_11 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3_11,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
```

#### Except bili, copper

```{r}
cox.train3_12 <- coxph(Surv(time,status) ~ albumin+protime+stage, data=pbc.train)
summary(cox.train3_12)
risk.train3_12 <- predict(object=cox.train3_12, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3_12 <- risk.train3_12
ROC.risk.train3_12 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3_12,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
```

#### Except bili, copper, albumin

```{r}
cox.train3_13 <- coxph(Surv(time,status) ~ protime+stage, data=pbc.train)
summary(cox.train3_13)
risk.train3_13 <- predict(object=cox.train3_13, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3_13 <- risk.train3_13
ROC.risk.train3_13 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3_13,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
```

#### Except bili, copper, albumin, stage

```{r}
cox.train3_14 <- coxph(Surv(time,status) ~ protime, data=pbc.train)
summary(cox.train3_14)
risk.train3_14 <- predict(object=cox.train3_14, newdata=pbc.test, type="risk") # risk score
pbc.test$risk3_14 <- risk.train3_14
ROC.risk.train3_14 <- timeROC(T=pbc.test$time,
                         delta=pbc.test$status,marker=pbc.test$risk3_14,
                         cause=1,weighting="marginal",
                         times=pbc.test$time)
```

#### Make AUC data

```{r}
dt2_1 <- data.frame(time=ROC.risk.train3[["times"]],
                    cox.train3=ROC.risk.train3[["AUC"]],
                    cox.train3_11=ROC.risk.train3_11[["AUC"]],
                    cox.train3_12=ROC.risk.train3_12[["AUC"]],
                    cox.train3_13=ROC.risk.train3_13[["AUC"]],
                    cox.train3_14=ROC.risk.train3_14[["AUC"]])
dt2_1.melt <- melt(data = dt2_1, 
                   id.vars = "time", 
                   measure.vars = c("cox.train3","cox.train3_11","cox.train3_12",
                                    "cox.train3_13","cox.train3_14"))
dt2_1.melt <- rename(dt2_1.melt,
                     replace=c("variable"="model",
                               "value"="AUC"))

time.density <- density(ROC.risk.train3[["times"]])
data1 <- data.frame(time=time.density[["x"]], density=time.density[["y"]])
data1$density.adjusted <- (max(na.omit(ROC.risk.train3_13[["AUC"]]))-min(na.omit(ROC.risk.train3_13[["AUC"]])))*(time.density[["y"]]-min(time.density[["y"]]))/(max(time.density[["y"]])-min(time.density[["y"]]))+min(na.omit(ROC.risk.train3_13[["AUC"]]))
```

#### Plot of Time dependent AUC with Adjusted Density of time and Status of censoring (Using ggplot2) (cox.train3~3_14)

```{r}
pbc$status1 <- ifelse(pbc$status==1,1,0.25)
ggplot() +
  geom_line(data=dt2_1.melt, aes(x=time, y=AUC,
                            group=model, color=model)) +
  geom_point(data=pbc, aes(x=time, y=status1, color=as.factor(status))) +
  geom_line(data=data1, aes(x=time, y=density.adjusted), size=1, colour='black') +
  scale_x_continuous(limits = c(0, max(pbc$time))) +
  labs(title="Time dependent AUC with risk score (cox.train3~3_14)",
       xlab="time", ylab="AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.train3~3_14)

```{r}
mean(ROC.risk.train3[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.train3_11[["AUC"]],na.rm=TRUE) #Except bili
mean(ROC.risk.train3_12[["AUC"]],na.rm=TRUE) #Except bili, copper
mean(ROC.risk.train3_13[["AUC"]],na.rm=TRUE) #Except bili, copper, albumin
mean(ROC.risk.train3_14[["AUC"]],na.rm=TRUE) #Except bili, copper, albumin, stage
```



