---
title: "Homework(9,10 weekend)"
author: "Hwang Seong-Yun"
date: '2020 11 3 '
output: html_document
---

## 8.7) Convert the covariance matrix S in Exercise 8.6 to a sample correlation matrix R. 

```{r}
library(tidyverse)
ex1 <- tibble(x1=c(108.28,152.36,95.04,65.45,62.97,263.99,265.19,285.06,92.01,165.68),x2=c(17.05,16.59,10.91,14.14,9.52,25.33,18.54,15.73,8.10,11.13))
```

### (a) Find the sample principal components y1_hat, y2_hat and their variances.

```{r}
(x_bar <- colMeans(ex1))
(S <- var(ex1))
(R <- cor(ex1))
lambda_R <- eigen(R)
lambda_R
(z1 <- (ex1$x1-mean(ex1$x1))/sd(ex1$x1))
(z2 <- (ex1$x2-mean(ex1$x2))/sd(ex1$x2))
y1_hat <- 0.7071068*z1+0.7071068*z2
y2_hat <- -0.7071068*z1+0.7071068*z2
var(y1_hat)
var(y2_hat)
```

#### y1_hat = 0.707 * z1 + 0.707 * z2, y2_hat = -0.707 * z1 + 0.707 * z2 
#### var(y1_hat)=1.686, var(y2_hat)=0.314

### (b) Compute the proportion of the total sample variance explained by y1_hat.

```{r}
lambda_R$values[[1]]/(lambda_R$values[[1]]+lambda_R$values[[2]])
```

#### 84.309 %

### (c) Compute the correlation coefficients r(y1_hat,zk) (k=1,2). 

```{r}
cor(y1_hat,z1)
cor(y1_hat,z2)
```

### (d) Compare the components obtained in Part a with those obtained in Exercise 8.6.

```{r}
lambda_S <- eigen(S)
lambda_S
y1_hat_S <- -0.99917338*ex1$x1-0.04065165*ex1$x2
y2_hat_S <- -0.04065165*ex1$x1-0.99917338*ex1$x2
var(y1_hat_S)
var(y2_hat_S)
```

#### Variable x1 and x2 has same unit. I think both methods are useful, but I want to take using sample covariance matrix S. Because, compared with R, it is better for explain characteristics of original data.

## 8.22) Consider the data on bulls in Table 1.10. Utilizing the seven variables SalePr, YrHgt, FtFrBody, PrctFFB, BkFat, SaleHt, SaleWt a principal component analysis using the covariance matrix S and the correlation matrix R.

```{r}
ex2 <- read.csv("C:/Users/HSY/Desktop/problem8_22.csv",sep=",",header=T)
head(ex2,10)
(S <- var(ex2))
(R <- cor(ex2))
```

### Using the covariance matrix S

```{r}
library(psych)
lambda_S <- eigen(S)
lambda_S
m_ex2 <- as.matrix(ex2)
X <- m_ex2%*%lambda_S$vectors
colnames(X) <- c("PC1","PC2","PC3","PC4","PC5","PC6","PC7")
library(ggplot2)
x <- as.data.frame(lambda_S$values)
x$PC <- 1:7
ggplot(data=x, aes(x=PC,y=lambda_S$values)) + 
  geom_line() + 
  geom_point(size=3) +
  ggtitle("Scree plot using the covariance matrix S")
```

#### Refer to the scree plot, appropriate number of components is 2.

```{r}
X <- as.data.frame(X)
ggplot(X, aes(x=PC1, y=PC2))+
  geom_point()+
  stat_ellipse() +
  ggtitle("PC1 versus PC2 using the covariance matrix S")
```

#### Refer to the scatter plot, some outliers are detected.

```{r}
qqnorm(X$PC1) ; qqline(X$PC1)
shapiro.test(X$PC1)
```

#### It seems that PC1 is not normality.

### Using the covariance matrix R

```{r}
lambda_R <- eigen(R)
lambda_R
m_ex2 <- as.matrix(ex2)
X1 <- m_ex2%*%lambda_R$vectors
colnames(X1) <- c("PC1","PC2","PC3","PC4","PC5","PC6","PC7")
x1 <- as.data.frame(lambda_R$values)
x1$PC <- 1:7
ggplot(data=x1, aes(x=PC,y=lambda_R$values)) + 
  geom_line() + 
  geom_point(size=3) +
  ggtitle("Scree plot using the covariance matrix R")
```

#### Refer to the scree plot, appropriate number of components is 4.

```{r}
X1 <- as.data.frame(X1)
ggplot(X1, aes(x=PC1, y=PC2))+
  geom_point()+
  stat_ellipse() +
  ggtitle("PC1 versus PC2 using the covariance matrix R")
```

#### Refer to the scatter plot, some outliers are detected.

```{r}
qqnorm(X1$PC1) ; qqline(X1$PC1)
shapiro.test(X1$PC1)
```

#### It seems that PC1 is not normality.

