---
title: "Time dependent AUC with survival data _ correct5"
author: "Hwang Seong-Yun"
date: '2022 4 29 '
output: html_document
---

# Loading packages

```{r warning = FALSE, message = FALSE}
library(survival)
library(timeROC)
library(timereg)
library(ggplot2)
library(reshape)
library(plyr)
library(tidyverse)
library(coxed)
library(gridExtra)
```

# Generating simulation data from Cox Regression Model 

## Background

- reference 1 : https://doi.org/10.5351/KJAS.2018.31.6.761
- reference 2 : https://doi.org/10.1017/psrm.2018.19


`-` real survival time $Y$ : Exponential distribution

`-` censoring time $C$ : Uniform distribution

- $h(t|x)=h_{0}(t)exp(f(x))$
- $H(t|x)=\int_{0}^{t} h(s|x)ds= \int_{0}^{t} h_{0}(s)exp(f(x))ds=H_{0}(t)exp(f(x))$
- $h(t|x)=\frac {f(t|x)}{S(t|x)}= -\frac {d}{dt} log[S(t|x)]$
- $log[S(t|x)]= -\int_{0}^{t} h(s|x)ds= -H(t|x)= -H_{0}(t)exp(f(x))$
- $S(t|x)=exp[-H_{0}(t)exp(f(x))]$

- In general,...
- $F(t|x)=1-exp[-H_{0}(t) exp(f(x))]$
- -> $F^{-1}(t|x)=H_{0}^{-1}[-log(1-t) exp(-f(x))]$
- By Inverse Transform Sampling, suppose that $U_{1}, U_{2} \overset{i.i.d}{\sim}\ U(0,1)$, then we can generate real survival time $Y$ from Cox Regression Model using this equation.
- $Y=H_{Y0}^{-1}[-log(U_{1})exp(-f(x))]$
- And by assuming non-informative censoring, we can generate censoring time $C$ like this equation. (Set $Y$ and $C$ are independent each other.)
- $C=H_{C0}^{-1}[-log(U_{2})]$
- And set censoring indicator $\delta=I(Y \leq C)$ and observed survival time $T=min(Y,C)$
- $H_{Y0}(t)$ : Basis cumulative hazard function of real survival time $Y$
- $H_{C0}(t)$ : Basis cumulative hazard function of censoring time $C$

- When parametric basis distribution of real survival time $Y$ is Exponential distribution and parametric basis distribution of censoring time $C$ is Uniform distribution ($Y$ and $C$ are independent each other), then ~
- $Y$ : $f_{Y0}(t)=\lambda e^{-\lambda t}, t>0$ -> $S_{Y0}(t)=e^{-\lambda t}$ -> $h_{Y0}(t)=\lambda$
- $C$ : $f_{C0}(t)=\frac {1}{\lambda}, 0<t<\lambda$ -> $S_{C0}(t)=1-\frac {t}{\lambda}$ -> $h_{C0}(t)=\frac {1}{\lambda -t}$
- $\therefore Y=-\frac{1}{\lambda_{1}} log(U_{1}) exp(-f(x))$, $C=\lambda_{2}(1-U_{2})$



## Generating simulation data

## Assume linear relationship

```{r}
# Assume parametric basis distribution is Exponential distribution... 
lambda1 <- 0.001 # parameter for generate Y  
lambda2 <- 1.5/lambda1 # parameter for generate C 
beta1 <- 3 # true coefficients
Y <- c() # real survival time
C <- c() # censoring time
t <- c() # observed survival time
censor <- c() # censoring indicator
X <- c() # explanatory variable
U1 <- c() # random variable from U(0,1) for generate Y
U2 <- c() # random variable from U(0,1) for generate C 
X2 <- c() # X^2

for (i in 1:200) {
  U1[i] <- runif(1)
  U2[i] <- runif(1)
  X[i] <- runif(1,min=0,max=1)
  X2[i] <- X[i]^2
  Y[i] <- -(1/lambda1)*log(U1[i])*exp(-beta1*X[i])
  C[i] <- (lambda2)*(1-U2[i])
  t[i] <- min(Y[i],C[i])
  censor[i] <- ifelse(Y[i]<=C[i],1,0)
}

mean(1-censor) # censoring rate
dt3 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
head(dt3,10)
min(dt3$time[dt3$censor==1])
```


### Time dependent AUC

### True model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1} x$
- True coefficients : $\beta_{1}=3$ 

### Set linear model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x$

```{r}
cox.lin.dt3 <- coxph(Surv(time,censor) ~ X, data=dt3)
summary(cox.lin.dt3)
risk.lin.dt3 <- predict(object=cox.lin.dt3, newdata=dt3, type="risk") # risk score
dt3$risk.lin.dt3 <- risk.lin.dt3
ROC.risk.lin.dt3 <- timeROC(T=dt3$time,
                         delta=dt3$censor,marker=dt3$risk.lin.dt3,
                         cause=1,weighting="marginal",
                         times=dt3$time)
ROC.risk.lin.dt3
```

### Set quadratic model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x+\beta_{2}x^2$

```{r}
cox.quad.dt3 <- coxph(Surv(time,censor) ~ X+X2, data=dt3)
summary(cox.quad.dt3)
risk.quad.dt3 <- predict(object=cox.quad.dt3, newdata=dt3, type="risk") # risk score
dt3$risk.quad.dt3 <- risk.quad.dt3
ROC.risk.quad.dt3 <- timeROC(T=dt3$time,
                         delta=dt3$censor,marker=dt3$risk.quad.dt3,
                         cause=1,weighting="marginal",
                         times=dt3$time)
ROC.risk.quad.dt3
```

#### Make AUC data

```{r}
# uncensored
uncensored.density3 <- density(dt3$time[dt3$censor==1])
data3_1 <- data.frame(time=uncensored.density3[["x"]], 
                    density=uncensored.density3[["y"]])
data3_1$density.adjusted <- (max(na.omit(ROC.risk.lin.dt3[["AUC"]]))-min(na.omit(ROC.risk.lin.dt3[["AUC"]])))*(uncensored.density3[["y"]]-min(uncensored.density3[["y"]]))/(max(uncensored.density3[["y"]])-min(uncensored.density3[["y"]]))+min(na.omit(ROC.risk.lin.dt3[["AUC"]]))

# censored
censored.density3 <- density(dt3$time[dt3$censor==0])
data3_2 <- data.frame(time=censored.density3[["x"]], 
                    density=censored.density3[["y"]])
data3_2$density.adjusted <- (max(na.omit(ROC.risk.lin.dt3[["AUC"]]))-min(na.omit(ROC.risk.lin.dt3[["AUC"]])))*(censored.density3[["y"]]-min(censored.density3[["y"]]))/(max(censored.density3[["y"]])-min(censored.density3[["y"]]))+min(na.omit(ROC.risk.lin.dt3[["AUC"]]))

res3 <- data.frame(time=ROC.risk.lin.dt3[["times"]],
                    cox.lin.dt3=ROC.risk.lin.dt3[["AUC"]],
                    cox.quad.dt3=ROC.risk.quad.dt3[["AUC"]])
res3 <- melt(data = res3, 
                   id.vars = "time", 
                   measure.vars = c("cox.lin.dt3","cox.quad.dt3"))
```

#### Time dependent AUC plot with ggplot2 (cox.lin.dt3, cox.quad.dt3)

```{r}
dt3$censor1 <- ifelse(dt3$censor==1,
                    1,min(na.omit(ROC.risk.lin.dt3[["AUC"]])))
ggplot() +
  geom_line(data=res3, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt3, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data3_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data3_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, max(dt3$time))) +
  geom_vline(xintercept = min(dt3$time[dt3$censor==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt3, cox.quad.dt3)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot() +
  geom_line(data=res3, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt3, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data3_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data3_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, min(dt3$time[dt3$censor==1]+50))) +
  scale_y_continuous(limits = c(min(na.omit(ROC.risk.lin.dt3[["AUC"]])), 1)) +
  geom_vline(xintercept = min(dt3$time[dt3$censor==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt3, cox.quad.dt3) \n (with minimum no censoring time point)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```


#### Mean value of Time dependent AUC (cox.lin.dt3, cox.quad.dt3)

```{r}
mean(ROC.risk.lin.dt3[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.quad.dt3[["AUC"]],na.rm=TRUE) 
```

#### Counting censored data

```{r}
dt3.censor <- dt3[dt3$censor==0,]
dt3.censor <- dt3.censor %>%
  mutate(n.cen=cut(dt3.censor$time, 
                   breaks=seq(from=0,to=1500,by=50)))
table(dt3.censor$n.cen)
```



## Assume quadratic relationship

```{r}
# Assume parametric basis distribution is Exponential distribution... 
lambda1 <- 0.001 # parameter for generate Y  
lambda2 <- 1.5/lambda1 # parameter for generate C 
beta1 <- 3 ; beta2 <- -2 # true coefficients
Y <- c() # real survival time
C <- c() # censoring time
t <- c() # observed survival time
censor <- c() # censoring indicator
X <- c() # explanatory variable
U1 <- c() # random variable from U(0,1) for generate Y
U2 <- c() # random variable from U(0,1) for generate C 
X2 <- c() # X^2

for (i in 1:200) {
  U1[i] <- runif(1)
  U2[i] <- runif(1)
  X[i] <- rnorm(1,mean=0,sd=1)
  X2[i] <- X[i]^2
  Y[i] <- -(1/lambda1)*log(U1[i])*exp(-beta1*X[i]-beta2*X2[i])
  C[i] <- (lambda2)*(1-U2[i])
  t[i] <- min(Y[i],C[i])
  censor[i] <- ifelse(Y[i]<=C[i],1,0)
}

mean(1-censor) # censoring rate
dt4 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
head(dt4,10)
min(dt4$time[dt4$censor==1])
```


### Time dependent AUC

### True model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1} x$
- True coefficients : $\beta_{1}=3, \beta_{2}=-2$ 

### Set linear model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x$

```{r}
cox.lin.dt4 <- coxph(Surv(time,censor) ~ X, data=dt4)
summary(cox.lin.dt4)
risk.lin.dt4 <- predict(object=cox.lin.dt4, newdata=dt4, type="risk") # risk score
dt4$risk.lin.dt4 <- risk.lin.dt4
ROC.risk.lin.dt4 <- timeROC(T=dt4$time,
                         delta=dt4$censor,marker=dt4$risk.lin.dt4,
                         cause=1,weighting="marginal",
                         times=dt4$time)
ROC.risk.lin.dt4
```

### Set quadratic model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x+\beta_{2}x^2$

```{r}
cox.quad.dt4 <- coxph(Surv(time,censor) ~ X+X2, data=dt4)
summary(cox.quad.dt4)
risk.quad.dt4 <- predict(object=cox.quad.dt4, newdata=dt4, type="risk") # risk score
dt4$risk.quad.dt4 <- risk.quad.dt4
ROC.risk.quad.dt4 <- timeROC(T=dt4$time,
                         delta=dt4$censor,marker=dt4$risk.quad.dt4,
                         cause=1,weighting="marginal",
                         times=dt4$time)
ROC.risk.quad.dt4
```

#### Make AUC data

```{r}
# uncensored
uncensored.density4 <- density(dt4$time[dt4$censor==1])
data4_1 <- data.frame(time=uncensored.density4[["x"]], 
                    density=uncensored.density4[["y"]])
data4_1$density.adjusted <- (max(na.omit(ROC.risk.lin.dt4[["AUC"]]))-min(na.omit(ROC.risk.lin.dt4[["AUC"]])))*(uncensored.density4[["y"]]-min(uncensored.density4[["y"]]))/(max(uncensored.density4[["y"]])-min(uncensored.density4[["y"]]))+min(na.omit(ROC.risk.lin.dt4[["AUC"]]))

# censored
censored.density4 <- density(dt4$time[dt4$censor==0])
data4_2 <- data.frame(time=censored.density4[["x"]], 
                    density=censored.density4[["y"]])
data4_2$density.adjusted <- (max(na.omit(ROC.risk.lin.dt4[["AUC"]]))-min(na.omit(ROC.risk.lin.dt4[["AUC"]])))*(censored.density4[["y"]]-min(censored.density4[["y"]]))/(max(censored.density4[["y"]])-min(censored.density4[["y"]]))+min(na.omit(ROC.risk.lin.dt4[["AUC"]]))

res4 <- data.frame(time=ROC.risk.lin.dt4[["times"]],
                    cox.lin.dt4=ROC.risk.lin.dt4[["AUC"]],
                    cox.quad.dt4=ROC.risk.quad.dt4[["AUC"]])
res4 <- melt(data = res4, 
                   id.vars = "time", 
                   measure.vars = c("cox.lin.dt4","cox.quad.dt4"))
```

#### Time dependent AUC plot with ggplot2 (cox.lin.dt4, cox.quad.dt4)

```{r}
dt4$censor1 <- ifelse(dt4$censor==1,
                    1,min(na.omit(ROC.risk.lin.dt4[["AUC"]])))
ggplot() +
  geom_line(data=res4, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt4, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data4_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data4_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, max(dt4$time))) +
  geom_vline(xintercept = min(dt4$time[dt4$censor==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt4, cox.quad.dt4)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot() +
  geom_line(data=res4, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt4, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data4_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data4_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, min(dt4$time[dt4$censor==1])+50)) +
  scale_y_continuous(limits = c(min(na.omit(ROC.risk.lin.dt4[["AUC"]])), 1)) +
  geom_vline(xintercept = min(dt4$time[dt4$censor==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt4, cox.quad.dt4) \n (with minimum no censoring time point)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.lin.dt4, cox.quad.dt4)

```{r}
mean(ROC.risk.lin.dt4[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.quad.dt4[["AUC"]],na.rm=TRUE) 
```

#### Counting censored data

```{r}
dt4.censor <- dt4[dt4$censor==0,]
dt4.censor <- dt4.censor %>%
  mutate(n.cen=cut(dt4.censor$time, 
                   breaks=seq(from=0,to=1500,by=50)))
table(dt4.censor$n.cen)
```



## Compare two results

```{r}
a <- ggplot()

fig1 <- a +
  geom_line(data=res3, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt3, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data3_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data3_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, 1500)) +
  geom_vline(xintercept = min(dt3$time[dt3$censor==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Assume linear relationship") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))

fig2 <- a +
  geom_line(data=res4, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt4, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data4_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data4_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, 1500)) +
  geom_vline(xintercept = min(dt4$time[dt4$censor==1]),
             linetype="dotted", color = "blue", size=1) +
  ggtitle("Assume quadratic relationship") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(fig1, fig2, nrow=2, ncol=1)
```



## simulation (repeat 100 times / time VS mean of AUC)

### Assume linear relationship

```{r}
lin.AUC <- matrix(nrow=length(seq(0,1500,0.2)),ncol=100)
quad.AUC <- matrix(nrow=length(seq(0,1500,0.2)),ncol=100)

for (i in 1:100) {
  
  # Assume parametric basis distribution is Exponential distribution... 
  lambda1 <- 0.001 # parameter for generate Y  
  lambda2 <- 1.5/lambda1 # parameter for generate C 
  beta1 <- 3 # true coefficients
  Y <- c() # real survival time
  C <- c() # censoring time
  t <- c() # observed survival time
  censor <- c() # censoring indicator
  X <- c() # explanatory variable
  U1 <- c() # random variable from U(0,1) for generate Y
  U2 <- c() # random variable from U(0,1) for generate C 
  X2 <- c() # X^2

  # Generate simulation data (Assume linear relationship)
  set.seed(i)
  for (j in 1:200) {
    U1[j] <- runif(1)
    U2[j] <- runif(1)
    X[j] <- runif(1,min=0,max=1)
    X2[j] <- X[j]^2
    Y[j] <- -(1/lambda1)*log(U1[j])*exp(-beta1*X[j])
    C[j] <- (lambda2)*(1-U2[j])
    t[j] <- min(Y[j],C[j])
    censor[j] <- ifelse(Y[j]<=C[j],1,0)
  }
  dt3 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
  
  # Set linear model
  cox.lin.dt3 <- coxph(Surv(time,censor) ~ X, data=dt3)
  risk.lin.dt3 <- predict(object=cox.lin.dt3, newdata=dt3, type="risk") 
  dt3$risk.lin.dt3 <- risk.lin.dt3
  ROC.risk.lin.dt3 <- timeROC(T=dt3$time,
                         delta=dt3$censor,marker=dt3$risk.lin.dt3,
                         cause=1,weighting="marginal",
                         times=seq(0,1500,0.2))
  
  # Set quadratic model
  cox.quad.dt3 <- coxph(Surv(time,censor) ~ X+X2, data=dt3)
  risk.quad.dt3 <- predict(object=cox.quad.dt3, newdata=dt3, type="risk") 
  dt3$risk.quad.dt3 <- risk.quad.dt3
  ROC.risk.quad.dt3 <- timeROC(T=dt3$time,
                         delta=dt3$censor,marker=dt3$risk.quad.dt3,
                         cause=1,weighting="marginal",
                         times=seq(0,1500,0.2))
  
  lin.AUC[,i] <- ROC.risk.lin.dt3[["AUC"]]
  quad.AUC[,i] <- ROC.risk.quad.dt3[["AUC"]]
  
  # print(i)
  
}

dat3 <- data.frame(time=seq(0,1500,0.2), lin.AUC=rowMeans(lin.AUC, na.rm=TRUE),
                   quad.AUC=rowMeans(quad.AUC, na.rm=TRUE))
plot(x=dat3$time, y=dat3$lin.AUC, xlab="time", ylab="mean of AUC",
     main="Assume linear relationship \n (Y : Exponential, C : Uniform)", type="l", col="red", ylim=c(0.5,1))
lines(x=dat3$time, y=dat3$quad.AUC, type="l", col="blue")
legend("bottomleft",c("lin.AUC","quad.AUC"),col=c("red","blue"),lty=c(1,1))
```

### Assume quadratic relationship

```{r}
lin.AUC <- matrix(nrow=length(seq(0,1500,0.2)),ncol=100)
quad.AUC <- matrix(nrow=length(seq(0,1500,0.2)),ncol=100)

for (i in 1:100) {
  
  # Assume parametric basis distribution is Exponential distribution... 
  lambda1 <- 0.001 # parameter for generate Y  
  lambda2 <- 1.5/lambda1 # parameter for generate C 
  beta1 <- 3 ; beta2 <- -2 # true coefficients
  Y <- c() # real survival time
  C <- c() # censoring time
  t <- c() # observed survival time
  censor <- c() # censoring indicator
  X <- c() # explanatory variable
  U1 <- c() # random variable from U(0,1) for generate Y
  U2 <- c() # random variable from U(0,1) for generate C 
  X2 <- c() # X^2

  # Generate simulation data (Assume quadratic relationship)
  set.seed(i)
  for (j in 1:200) {
    U1[j] <- runif(1)
    U2[j] <- runif(1)
    X[j] <- rnorm(1,mean=0,sd=1)
    X2[j] <- X[j]^2
    Y[j] <- -(1/lambda1)*log(U1[j])*exp(-beta1*X[j]-beta2*X2[j])
    C[j] <- (lambda2)*(1-U2[j])
    t[j] <- min(Y[j],C[j])
    censor[j] <- ifelse(Y[j]<=C[j],1,0)
  }
  dt4 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)

  # Set linear model
  cox.lin.dt4 <- coxph(Surv(time,censor) ~ X, data=dt4)
  risk.lin.dt4 <- predict(object=cox.lin.dt4, newdata=dt4, type="risk") 
  dt4$risk.lin.dt4 <- risk.lin.dt4
  ROC.risk.lin.dt4 <- timeROC(T=dt4$time,
                         delta=dt4$censor,marker=dt4$risk.lin.dt4,
                         cause=1,weighting="marginal",
                         times=seq(0,1500,0.2))
  
  # Set quadratic model
  cox.quad.dt4 <- coxph(Surv(time,censor) ~ X+X2, data=dt4)
  risk.quad.dt4 <- predict(object=cox.quad.dt4, newdata=dt4, type="risk") 
  dt4$risk.quad.dt4 <- risk.quad.dt4
  ROC.risk.quad.dt4 <- timeROC(T=dt4$time,
                         delta=dt4$censor,marker=dt4$risk.quad.dt4,
                         cause=1,weighting="marginal",
                         times=seq(0,1500,0.2))

  lin.AUC[,i] <- ROC.risk.lin.dt4[["AUC"]]
  quad.AUC[,i] <- ROC.risk.quad.dt4[["AUC"]]
  
  # print(i)
  
}

dat4 <- data.frame(time=seq(0,1500,0.2), lin.AUC=rowMeans(lin.AUC, na.rm=TRUE),
                   quad.AUC=rowMeans(quad.AUC, na.rm=TRUE))
plot(x=dat4$time, y=dat4$lin.AUC, xlab="time", ylab="mean of AUC",
     main="Assume quadratic relationship \n (Y : Exponential, C : Uniform)", type="l", col="red", ylim=c(0.5,1))
lines(x=dat4$time, y=dat4$quad.AUC, type="l", col="blue")
legend("bottomleft",c("lin.AUC","quad.AUC"),col=c("red","blue"),lty=c(1,1))
```




# Generate survival data with time dependent covariates

## Background

- reference  : https://dx.doi.org/10.21037/atm.2018.02.12


`-` real survival time $Y$ : Exponential distribution

`-` censoring time $C$ : Uniform distribution

- $h(t|x)=h_{0}(t)exp[f(x(t))]$
- $H(t|x)=\int_{0}^{t} h(s|x)ds= \int_{0}^{t} h_{0}(s)exp[f(x(t))]ds=H_{0}(t)exp[f(x(t))]$
- $h(t|x)=\frac {f(t|x)}{S(t|x)}= -\frac {d}{dt} log[S(t|x)]$
- $log[S(t|x)]= -\int_{0}^{t} h(s|x)ds= -H(t|x)= -H_{0}(t)exp[f(x(t))]$
- $S(t|x)=exp[-H_{0}(t)exp[f(x(t))]]$

- In general,...
- $F(t|x)=1-exp[-H_{0}(t) exp[f(x(t))]]$
- -> $F^{-1}(t|x)=H_{0}^{-1}[-log(1-t) exp[-f(x(t))]]$
- By Inverse Transform Sampling, suppose that $U_{1}, U_{2} \overset{i.i.d}{\sim}\ U(0,1)$, then we can generate real survival time $Y$ from Cox Regression Model using this equation.
- $Y=H_{Y0}^{-1}[-log(U_{1})exp[-f(x(t))]]$
- And by assuming non-informative censoring, we can generate censoring time $C$ like this equation. (Set $Y$ and $C$ are independent each other.)
- $C=H_{C0}^{-1}[-log(U_{2})]$
- And set censoring indicator $\delta=I(Y \leq C)$ and observed survival time $T=min(Y,C)$
- $H_{Y0}(t)$ : Basis cumulative hazard function of real survival time $Y$
- $H_{C0}(t)$ : Basis cumulative hazard function of censoring time $C$

- When parametric basis distribution of real survival time $Y$ is Exponential distribution and parametric basis distribution of censoring time $C$ is Uniform distribution ($Y$ and $C$ are independent each other), then ~
- $Y$ : $f_{Y0}(t)=\lambda e^{-\lambda t}, t>0$ -> $S_{Y0}(t)=e^{-\lambda t}$ -> $h_{Y0}(t)=\lambda$
- $C$ : $f_{C0}(t)=\frac {1}{\lambda}, 0<t<\lambda$ -> $S_{C0}(t)=1-\frac {t}{\lambda}$ -> $h_{C0}(t)=\frac {1}{\lambda -t}$
- $\therefore Y=-\frac{1}{\lambda_{1}} log(U_{1}) exp[-f(x(t))]$, $C=\lambda_{2}(1-U_{2})$

- In this simulation, I take this Cox regression model with time dependent covariates.
- $h(t|x)=h_{0}(t)exp[\beta_{1}x+\beta_{2}g(x,t)]$
- $g(x,t)=\alpha_{0}+\alpha_{1}x+\alpha_{2}xt$


## Generating simulation data

```{r}
# Assume parametric basis distribution is Exponential distribution... 
lambda1 <- 0.001 # parameter for generate Y  
lambda2 <- 1.5/lambda1 # parameter for generate C 
beta1 <- 3 ; beta2 <- -2 # true coefficients 1
alpha0 <- 1 ; alpha1 <- 1 ; alpha2 <- -3 # true coefficients 2
Y <- c() # real survival time
C <- c() # censoring time
t <- c() # observed survival time
censor <- c() # censoring indicator
X <- c() # explanatory variable
U1 <- c() # random variable from U(0,1) for generate Y
U2 <- c() # random variable from U(0,1) for generate C 
X2 <- c() # X^2

for (i in 1:200) {
  U1[i] <- runif(1)
  U2[i] <- runif(1)
  X[i] <- rnorm(1,mean=0,sd=1)
  X2[i] <- X[i]^2
  Y[i] <- -(1/lambda1)*log(U1[i])*exp(-beta1*X[i])
  C[i] <- (lambda2)*(1-U2[i])
  t[i] <- min(Y[i],C[i])
  censor[i] <- ifelse(Y[i]<=C[i],1,0)
}

mean(1-censor) # censoring rate
dt3 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
head(dt3,10)
min(dt3$time[dt3$censor==1])
```


