---
title: "Time dependent AUC with survival data _ correct1"
author: "Hwang Seong-Yun"
date: '2022 4 11 '
output: html_document
---

# Loading packages

```{r}
library(survival)
library(timeROC)
library(timereg)
library(ggplot2)
library(reshape)
library(plyr)
library(tidyverse)
library(coxed)
library(gridExtra)
```

# Generating simulation data from Cox regression 

## Background

- reference 1 : https://doi.org/10.5351/KJAS.2018.31.6.761
- reference 2 : https://doi.org/10.1017/psrm.2018.19

- $h(t|x)=h_{0}(t)exp(f(x))$
- $H(t|x)=\int_{0}^{t} h(s|x)ds= \int_{0}^{t} h_{0}(s)exp(f(x))ds=H_{0}(t)exp(f(x))$
- $h(t|x)=\frac {f(t|x)}{S(t|x)}= -\frac {d}{dt} log[S(t|x)]$
- $log[S(t|x)]= -\int_{0}^{t} h(s|x)ds= -H(t|x)= -H_{0}(t)exp(f(x))$
- $S(t|x)=exp[-H_{0}(t)exp(f(x))]$

- In general,...
- $F(t|x)=1-exp[-H_{0}(t) exp(f(x))]$
- -> $F^{-1}(t|x)=H_{0}^{-1}[-log(1-t) exp(-f(x))]$
- By Inverse Transform Sampling, suppose that $U_{1}, U_{2} \overset{i.i.d}{\sim}\ U(0,1)$, then we can generate real survival time $Y$ from Cox Proportional Hazard Model using this equation.
- $Y=H_{Y0}^{-1}[-log(U_{1})exp(-f(x))]$
- And by assuming non-informative censoring, we can generate censoring time $C$ like this equation. (Set $Y$ and $C$ are independent each other.)
- $C=H_{C0}^{-1}[-log(U_{2})]$
- And set censoring indicator $\delta=I(Y \leq C)$ and observed survival time $T=min(Y,C)$
- $H_{Y0}(t|x)$ : Basis cumulative hazard function of real survival time $Y$
- $H_{C0}(t)$ : Basis cumulative hazard function of censoring time $C$

- When parametric basis distribution of real survival time $Y$ and censoring time $C$ are Exponential distribution and independent each other, then ~
- $f_{0}(t)=\lambda e^{-\lambda t}, t>0$ -> $S_{0}(t)=e^{-\lambda t}$ -> $h_{0}(t)=\lambda$
- $\therefore Y=-\frac{1}{\lambda} log(U_{1}) exp(-f(x))$, $C=-\frac{1}{\lambda} log(U_{2})$






# Generating simulation data


## Assume linear relationship

```{r}
# Assume parametric basis distribution is Exponential distribution... 
lambda1 <- 0.1 # parameter for generate Y  
lambda2 <- 0.05 # parameter for generate C 
beta1 <- 3 # true coefficients
Y <- c() # real survival time
C <- c() # censoring time
t <- c() # observed survival time
censor <- c() # censoring indicator
X <- c() # explanatory variable
U1 <- c() # random variable from U(0,1) for generate Y
U2 <- c() # random variable from U(0,1) for generate C 
X2 <- c() # X^2

for (i in 1:10000) {
  U1[i] <- runif(1)
  U2[i] <- runif(1)
  # X[i] <- runif(1,min=0,max=10)
  X[i] <- rnorm(1,mean=0,sd=1)
  X2[i] <- X[i]^2
  Y[i] <- -(1/lambda1)*log(U1[i])*exp(-beta1*X[i])
  C[i] <- -(1/lambda2)*log(U2[i])
  t[i] <- min(Y[i],C[i])
  censor[i] <- ifelse(Y[i]<=C[i],1,0)
}

mean(1-censor) # censoring rate
dt1 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
head(dt1,20)
# dt1[c(order(dt1$time)),]$censor
```


### Time dependent AUC

### True model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1} x$
- True coefficients : $\beta_{1}=3$ 

### Set linear model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x$

```{r}
cox.lin.dt1 <- coxph(Surv(time,censor) ~ X, data=dt1)
summary(cox.lin.dt1)
risk.lin.dt1 <- predict(object=cox.lin.dt1, newdata=dt1, type="risk") # risk score
dt1$risk.lin.dt1 <- risk.lin.dt1
ROC.risk.lin.dt1 <- timeROC(T=dt1$time,
                         delta=dt1$censor,marker=dt1$risk.lin.dt1,
                         cause=1,weighting="marginal",
                         times=dt1$time)
ROC.risk.lin.dt1
```

### Set quadratic model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x+\beta_{2}x^2$

```{r}
cox.quad.dt1 <- coxph(Surv(time,censor) ~ X+X2, data=dt1)
summary(cox.quad.dt1)
risk.quad.dt1 <- predict(object=cox.quad.dt1, newdata=dt1, type="risk") # risk score
dt1$risk.quad.dt1 <- risk.quad.dt1
ROC.risk.quad.dt1 <- timeROC(T=dt1$time,
                         delta=dt1$censor,marker=dt1$risk.quad.dt1,
                         cause=1,weighting="marginal",
                         times=dt1$time)
ROC.risk.quad.dt1
```

#### Make AUC data

```{r}
# uncensored
uncensored.density1 <- density(dt1$time[dt1$censor==1])
data1_1 <- data.frame(time=uncensored.density1[["x"]], 
                    density=uncensored.density1[["y"]])
data1_1$density.adjusted <- (max(na.omit(ROC.risk.lin.dt1[["AUC"]]))-min(na.omit(ROC.risk.lin.dt1[["AUC"]])))*(uncensored.density1[["y"]]-min(uncensored.density1[["y"]]))/(max(uncensored.density1[["y"]])-min(uncensored.density1[["y"]]))+min(na.omit(ROC.risk.lin.dt1[["AUC"]]))

# censored
censored.density1 <- density(dt1$time[dt1$censor==0])
data1_2 <- data.frame(time=censored.density1[["x"]], 
                    density=censored.density1[["y"]])
data1_2$density.adjusted <- (max(na.omit(ROC.risk.lin.dt1[["AUC"]]))-min(na.omit(ROC.risk.lin.dt1[["AUC"]])))*(censored.density1[["y"]]-min(censored.density1[["y"]]))/(max(censored.density1[["y"]])-min(censored.density1[["y"]]))+min(na.omit(ROC.risk.lin.dt1[["AUC"]]))

res1 <- data.frame(time=ROC.risk.lin.dt1[["times"]],
                    cox.lin.dt1=ROC.risk.lin.dt1[["AUC"]],
                    cox.quad.dt1=ROC.risk.quad.dt1[["AUC"]])
res1 <- melt(data = res1, 
                   id.vars = "time", 
                   measure.vars = c("cox.lin.dt1","cox.quad.dt1"))
```

#### Time dependent AUC plot with ggplot2 (cox.lin.dt1, cox.quad.dt1)

```{r}
dt1$censor1 <- ifelse(dt1$censor==1,
                    1,min(na.omit(ROC.risk.lin.dt1[["AUC"]])))
ggplot() +
  geom_line(data=res1, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt1, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data1_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data1_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(min(dt1$time), max(dt1$time))) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt1, cox.quad.dt1)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.lin.dt1, cox.quad.dt1)

```{r}
mean(ROC.risk.lin.dt1[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.quad.dt1[["AUC"]],na.rm=TRUE) 
```

#### Counting censored data

```{r}
dt1.censor <- dt1[dt1$censor==0,]
dt1.censor <- dt1.censor %>%
  mutate(n.cen=cut(dt1.censor$time, 
                   breaks=seq(from=0,to=200,by=5)))
table(dt1.censor$n.cen)
```



## Assume quadratic relationship

```{r}
# Assume parametric basis distribution is Exponential distribution... 
lambda1 <- 0.1 # parameter for generate Y  
lambda2 <- 0.05 # parameter for generate C 
beta1 <- 3 ; beta2 <- 2 # true coefficients
Y <- c() # real survival time
C <- c() # censoring time
t <- c() # observed survival time
censor <- c() # censoring indicator
X <- c() # explanatory variable
U1 <- c() # random variable from U(0,1) for generate Y
U2 <- c() # random variable from U(0,1) for generate C 
X2 <- c() # X^2

for (i in 1:10000) {
  U1[i] <- runif(1)
  U2[i] <- runif(1)
  # X[i] <- runif(1,min=0,max=10)
  X[i] <- rnorm(1,mean=0,sd=1)
  X2[i] <- X[i]^2
  Y[i] <- -(1/lambda1)*log(U1[i])*exp(-beta1*X[i]-beta2*X2[i])
  C[i] <- -(1/lambda2)*log(U2[i])
  t[i] <- min(Y[i],C[i])
  censor[i] <- ifelse(Y[i]<=C[i],1,0)
}

mean(1-censor) # censoring rate
dt2 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
head(dt2,20)
# dt2[c(order(dt2$time)),]$censor
```


### Time dependent AUC

### True model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1} x + \beta_{2} x^2$
- True coefficients : $\beta_{1}=3, \beta_{2}=2$ 

### Set linear model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x$

```{r}
cox.lin.dt2 <- coxph(Surv(time,censor) ~ X, data=dt2)
summary(cox.lin.dt2)
risk.lin.dt2 <- predict(object=cox.lin.dt2, newdata=dt2, type="risk") # risk score
dt2$risk.lin.dt2 <- risk.lin.dt2
ROC.risk.lin.dt2 <- timeROC(T=dt2$time,
                         delta=dt2$censor,marker=dt2$risk.lin.dt2,
                         cause=1,weighting="marginal",
                         times=dt2$time)
ROC.risk.lin.dt2
```

### Set quadratic model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x+\beta_{2}x^2$

```{r}
cox.quad.dt2 <- coxph(Surv(time,censor) ~ X+X2, data=dt2)
summary(cox.quad.dt2)
risk.quad.dt2 <- predict(object=cox.quad.dt2, newdata=dt2, type="risk") # risk score
dt2$risk.quad.dt2 <- risk.quad.dt2
ROC.risk.quad.dt2 <- timeROC(T=dt2$time,
                         delta=dt2$censor,marker=dt2$risk.quad.dt2,
                         cause=1,weighting="marginal",
                         times=dt2$time)
ROC.risk.quad.dt2
```

#### Make AUC data

```{r}
# uncensored
uncensored.density2 <- density(dt2$time[dt2$censor==1])
data2_1 <- data.frame(time=uncensored.density2[["x"]], 
                    density=uncensored.density2[["y"]])
data2_1$density.adjusted <- (max(na.omit(ROC.risk.lin.dt2[["AUC"]]))-min(na.omit(ROC.risk.lin.dt2[["AUC"]])))*(uncensored.density2[["y"]]-min(uncensored.density2[["y"]]))/(max(uncensored.density2[["y"]])-min(uncensored.density2[["y"]]))+min(na.omit(ROC.risk.lin.dt2[["AUC"]]))

# censored
censored.density2 <- density(dt2$time[dt2$censor==0])
data2_2 <- data.frame(time=censored.density2[["x"]], 
                    density=censored.density2[["y"]])
data2_2$density.adjusted <- (max(na.omit(ROC.risk.lin.dt2[["AUC"]]))-min(na.omit(ROC.risk.lin.dt2[["AUC"]])))*(censored.density2[["y"]]-min(censored.density2[["y"]]))/(max(censored.density2[["y"]])-min(censored.density2[["y"]]))+min(na.omit(ROC.risk.lin.dt2[["AUC"]]))

res2 <- data.frame(time=ROC.risk.lin.dt2[["times"]],
                    cox.lin.dt2=ROC.risk.lin.dt2[["AUC"]],
                    cox.quad.dt2=ROC.risk.quad.dt2[["AUC"]])
res2 <- melt(data = res2, 
                   id.vars = "time", 
                   measure.vars = c("cox.lin.dt2","cox.quad.dt2"))
```

#### Time dependent AUC plot with ggplot2 (cox.lin.dt2, cox.quad.dt2)

```{r}
dt2$censor1 <- ifelse(dt2$censor==1,
                    1,min(na.omit(ROC.risk.lin.dt2[["AUC"]])))
ggplot() +
  geom_line(data=res2, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt2, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data2_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data2_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(min(dt2$time), max(dt2$time))) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt2, cox.quad.dt2)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.lin.dt2, cox.quad.dt2)

```{r}
mean(ROC.risk.lin.dt2[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.quad.dt2[["AUC"]],na.rm=TRUE) 
```

#### Counting censored data

```{r}
dt2.censor <- dt2[dt2$censor==0,]
dt2.censor <- dt2.censor %>%
  mutate(n.cen=cut(dt2.censor$time, 
                   breaks=seq(from=0,to=200,by=5)))
table(dt2.censor$n.cen)
```



## Compare two results

```{r}
a <- ggplot()

fig1 <- a +
  geom_line(data=res1, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt1, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data1_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data1_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(min(dt1$time), max(dt1$time))) +
  ggtitle("Assume linear relationship") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))

fig2 <- a +
  geom_line(data=res2, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt2, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data2_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data2_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(min(dt2$time), max(dt2$time))) +
  ggtitle("Assume quadratic relationship") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(fig1, fig2, nrow=2, ncol=1)
```


