---
title: "Cure model in Survival Analysis"
author: "Hwang Seong-Yun"
date: '2021 5 4 '
output: html_document
---

## Mixture cure model

### Loading packages
- Fit semiparametric PH(Proportional Hazard) mixture cure model or AFT(Accelerated Failure Time) mixture cure model by the EM algorithm

```{r}
library(smcure)
```

### Mixture cure model
- $S_{pop}(t|x,z)=[1-p(x)]+p(x)S_{u}(t|z)$  
- $S_{pop}(t|x,z)=P(T>t|X=x,Z=z)$ : The survival function 
- $p(x)=P(B=1|X=x)$ : The probability of being susceptible (often called the incidence of the model) 
- $S_{u}(t|z)=P(T>t|Z=z,B=1)$ : The (proper) conditional survival function of the susceptibles (often called the latency of the model)
- $B=I(T<\infty)$ : The latent binary variable indicating whether someone is cured or not 
- $F_{pop}(t|x,z)=p(x)F_{u}(t|z)$

### Fit PH(Proportional Hazard) mixture cure model

#### Eastern Cooperative Oncology Group (ECOG) data
- The melanoma data from the Eastern Cooperative Oncology Group (ECOG) phase III clinical trial e1684 which is used for modeling semicure PH mixture cure model (Kirkwood, et al., 1996)
- TRT : 0=control group, 1=IFN treatment group 
- FAILTIME : observed relapse-free time
- FAILCENS : censoring indicator, 1=event of interest happens, and 0=censoring
- AGE : continuous variable, which is centered to the mean
- SEX : 0 for male, 1 for female

```{r}
data(e1684)
```

#### Fit logistic/Cox PH mixture cure model
- $p(x)=[exp(\gamma^{T}x)]/[1+exp(\gamma^{T}x)]$
- $S_{u}(t|z)=S_{0}(t)^{exp(\beta^{T}z)}$

```{r}
# fit logistic/Cox PH mixture cure model
pd <- smcure(Surv(FAILTIME,FAILCENS)~TRT+SEX+AGE,cureform=~TRT+SEX+AGE,data=e1684,model="ph",link="logit",Var=FALSE)
pd
printsmcure(pd,Var=FALSE)
```

- Predicted population survival function can be calculated by the following equation
- ${\hat{S}}_{pop}(t|x,z)=[1-{\hat{p}}(x)]+{\hat{p}}(x){\hat{S}}_{u}(t|z)$

```{r}
# plot predicted survival curves for male with median centered age by treatment groups
predm <- predictsmcure(pd,newX=cbind(c(1,0,1),c(0,0,1),c(0.475,0.613,0.826)),newZ=cbind(c(1,0,1),c(0,0,1),c(0.475,0.613,0.826)),model="ph")
predm
plotpredictsmcure(predm,model="ph")
```

### Fit AFT(Accelerated Failure Time) mixture cure model

#### Bone marrow transplant study data
- Bone marrow transplant study for the refractory acute lymphoblastic leukemia patients which is widely used in the AFTMC model (Kersey, et al., 1987)
- Time : time to death
- Status : censoring indicator, 1=event of interest happens, and 0=censoring
- TRT : 1 for autologous treatment group; 0 for allogeneic treatment group

```{r}
data(bmt)
```

#### Fit logistic/semiparametric AFT models
- $log(T)=\beta^{T}Z+\epsilon$

```{r}
# fit AFT mixture cure model
bmtfit <- smcure(formula=Surv(Time,Status)~TRT,cureform=~TRT,data=bmt,model="aft",Var=FALSE)
printsmcure(bmtfit,Var=FALSE)
```

```{r}
# plot predicted Survival curves by treatment groups
predbmt=predictsmcure(bmtfit,newX=c(0,1),newZ=c(0,1),model="aft")
plotpredictsmcure(predbmt,model="aft")
```


## Promotion time cure model

### Loading packages
- Fits Semiparametric Promotion Time Cure Models, taking into account (using a corrected score approach or the SIMEX algorithm) or not the measurement error in the covariates, using a backfitting approach to maximize the likelihood. 

```{r}
library(miCoPTCM)
```

### Promotion time cure model
- $S_{pop}(t|x)=exp[-\theta(x)F(t)]$ 
- $S_{pop}(t|x)=P(T>t|X=x)$ : The survival function
- $F(t)$ : A proper baseline distribution function
- $\theta(x)$ : Captures the effect of the covariates $x$ on the survival function $S_{pop}(t|x)$ 
- $P(B=0|X=x)=exp[-\theta(x)]$ : The cure proportion 
- $W=X+U$ : The classical additive measurement error model ($W$ : The vector of observed covariates, $U$ : The vector of measurement errors($U \sim N_{p}(0,V)$, where $V$ is known, and $U$ is independent of $X$, $(T,C)$ and $W$ are independent given $X$))

### Corrected score approach
- Fits a Semiparametric Promotion Time Cure Model, taking into account (using a corrected score approach) or not the measurement error in the covariates, using a backfitting approach to maximize the likelihood. Both methods were introduced in Ma and Yin (2008).

```{r}
library("survival")
## Data generation
set.seed(123)
n <- 200
varCov <- matrix(nrow=3,ncol=3,0)
varCov[2,2] <- 0.1^1
X1 <- (runif(n)-0.5)/sqrt(1/12) 
V <- round(X1+rnorm(n,rep(0,3),varCov[2,2]),7)# covariate with measurement error
Xc <- round(as.numeric(runif(n)<0.5),7) # covariate without measurement error

 # censoring times: truncated exponential distribution
C <- round(rexp(n,1/5),5) 
Cbin <- (C>30)
while (sum(Cbin)>0) {
	C[Cbin] <- round(rexp(sum(Cbin),1/5),5)
	Cbin <- (C>30) }

expb <- exp(0.5+X1-0.5*Xc) 
cure <- exp(-expb) # cure probabilities

 # event times with baseline cdf of a truncated exponential 
U <- runif(n)
d <- rep(NA,n)
T <- round(-6*log(1+(1-exp(-20/6))*log(1-(1-cure)*U)/expb),5) 
T[(runif(n)<cure)] <- 99999 # cured subjects

Tobs <- rep(NA,n)
Tobs <- pmin(C,T) # observed times
Tmax <- max(Tobs[Tobs==T])
d <- (Tobs==T) # censoring indicator
	
Dat <- data.frame(Tobs,d,V,Xc)
#colnames(Dat) <- c("Tobs","d","V","Xc")

## Model estimation
fm <- formula(Surv(Tobs,d)~V+Xc)
resMY <- PTCMestimBF(fm,Dat,varCov=varCov,init=rnorm(3))
resMY
summary(resMY)
```

### SIMEX approach
- Fits a Semiparametric Promotion Time Cure Model with mismeasured covariates, using the SIMEX algorithm based on a backfitting procedure. This approach is introduced in Bertrand et al. (2015).

```{r}
## Not run: 
library("survival")
## Data generation
set.seed(123)
n <- 200
varCov <- matrix(nrow=3,ncol=3,0)
varCov[2,2] <- 0.1^1
X1 <- (runif(n)-0.5)/sqrt(1/12) 
V <- round(X1+rnorm(n,rep(0,3),varCov[2,2]),7)# covariate with measurement error
Xc <- round(as.numeric(runif(n)<0.5),7) # covariate without measurement error

 # censoring times: truncated exponential distribution
C <- round(rexp(n,1/5),5) 
Cbin <- (C>30)
while (sum(Cbin)>0) {
	C[Cbin] <- round(rexp(sum(Cbin),1/5),5)
	Cbin <- (C>30) }

expb <- exp(0.5+X1-0.5*Xc) 
cure <- exp(-expb) # cure probabilities

 # event times with baseline cdf of a truncated exponential 
U <- runif(n)
d <- rep(NA,n)
T <- round(-6*log(1+(1-exp(-20/6))*log(1-(1-cure)*U)/expb),5) 
T[(runif(n)<cure)] <- 99999 # cured subjects

Tobs <- rep(NA,n)
Tobs <- pmin(C,T) # observed times
Tmax <- max(Tobs[Tobs==T])
d <- (Tobs==T) # censoring indicator
	
Dat <- data.frame(Tobs,d,V,Xc)

## Model estimation
fm <- formula(Surv(Tobs,d)~V+Xc)
resSimex <- PTCMestimSIMEX(fm,Dat,errorDistEstim="normal", 
varCov=varCov,nBack=10000,eps=1e-8,Nu=c(0,0.5,1,1.5,2),B=50,init=rnorm(3),orderExtrap=1:3,multMaxTime=2)
resSimex
summary(resSimex)
## End(Not run)
```

