---
title: "Time dependent AUC with survival data _ correct3"
author: "Hwang Seong-Yun"
date: '2022 4 15 '
output: html_document
---

# Loading packages

```{r warning = FALSE, message = FALSE}
library(survival)
library(timeROC)
library(timereg)
library(ggplot2)
library(reshape)
library(plyr)
library(tidyverse)
library(coxed)
library(gridExtra)
```

# Generating simulation data from Cox Regression Model 

## Background

- reference 1 : https://doi.org/10.5351/KJAS.2018.31.6.761
- reference 2 : https://doi.org/10.1017/psrm.2018.19

- $h(t|x)=h_{0}(t)exp(f(x))$
- $H(t|x)=\int_{0}^{t} h(s|x)ds= \int_{0}^{t} h_{0}(s)exp(f(x))ds=H_{0}(t)exp(f(x))$
- $h(t|x)=\frac {f(t|x)}{S(t|x)}= -\frac {d}{dt} log[S(t|x)]$
- $log[S(t|x)]= -\int_{0}^{t} h(s|x)ds= -H(t|x)= -H_{0}(t)exp(f(x))$
- $S(t|x)=exp[-H_{0}(t)exp(f(x))]$

- In general,...
- $F(t|x)=1-exp[-H_{0}(t) exp(f(x))]$
- -> $F^{-1}(t|x)=H_{0}^{-1}[-log(1-t) exp(-f(x))]$
- By Inverse Transform Sampling, suppose that $U_{1}, U_{2} \overset{i.i.d}{\sim}\ U(0,1)$, then we can generate real survival time $Y$ from Cox Regression Model using this equation.
- $Y=H_{Y0}^{-1}[-log(U_{1})exp(-f(x))]$
- And by assuming non-informative censoring, we can generate censoring time $C$ like this equation. (Set $Y$ and $C$ are independent each other.)
- $C=H_{C0}^{-1}[-log(U_{2})]$
- And set censoring indicator $\delta=I(Y \leq C)$ and observed survival time $T=min(Y,C)$
- $H_{Y0}(t)$ : Basis cumulative hazard function of real survival time $Y$
- $H_{C0}(t)$ : Basis cumulative hazard function of censoring time $C$

- When parametric basis distribution of real survival time $Y$ and censoring time $C$ are Exponential distribution and independent each other, then ~
- $f_{0}(t)=\lambda e^{-\lambda t}, t>0$ -> $S_{0}(t)=e^{-\lambda t}$ -> $h_{0}(t)=\lambda$
- $\therefore Y=-\frac{1}{\lambda_{1}} log(U_{1}) exp(-f(x))$, $C=-\frac{1}{\lambda_{2}} log(U_{2})$






# Generating simulation data


## Assume linear relationship

```{r}
# Assume parametric basis distribution is Exponential distribution... 
lambda1 <- 0.01 # parameter for generate Y  
lambda2 <- 0.01 # parameter for generate C 
beta1 <- 3 # true coefficients
Y <- c() # real survival time
C <- c() # censoring time
t <- c() # observed survival time
censor <- c() # censoring indicator
X <- c() # explanatory variable
U1 <- c() # random variable from U(0,1) for generate Y
U2 <- c() # random variable from U(0,1) for generate C 
X2 <- c() # X^2

for (i in 1:1000) {
  U1[i] <- runif(1)
  U2[i] <- runif(1)
  # X[i] <- runif(1,min=0,max=10)
  X[i] <- rnorm(1,mean=0,sd=1)
  X2[i] <- X[i]^2
  Y[i] <- -(1/lambda1)*log(U1[i])*exp(-beta1*X[i])
  C[i] <- -(1/lambda2)*log(U2[i])
  t[i] <- min(Y[i],C[i])
  censor[i] <- ifelse(Y[i]<=C[i],1,0)
}

mean(1-censor) # censoring rate
dt1 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
head(dt1,20)
# dt1[c(order(dt1$time)),]$censor
```


### Time dependent AUC

### True model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1} x$
- True coefficients : $\beta_{1}=3$ 

### Set linear model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x$

```{r}
cox.lin.dt1 <- coxph(Surv(time,censor) ~ X, data=dt1)
summary(cox.lin.dt1)
risk.lin.dt1 <- predict(object=cox.lin.dt1, newdata=dt1, type="risk") # risk score
dt1$risk.lin.dt1 <- risk.lin.dt1
ROC.risk.lin.dt1 <- timeROC(T=dt1$time,
                         delta=dt1$censor,marker=dt1$risk.lin.dt1,
                         cause=1,weighting="marginal",
                         times=dt1$time)
ROC.risk.lin.dt1
```

### Set quadratic model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x+\beta_{2}x^2$

```{r}
cox.quad.dt1 <- coxph(Surv(time,censor) ~ X+X2, data=dt1)
summary(cox.quad.dt1)
risk.quad.dt1 <- predict(object=cox.quad.dt1, newdata=dt1, type="risk") # risk score
dt1$risk.quad.dt1 <- risk.quad.dt1
ROC.risk.quad.dt1 <- timeROC(T=dt1$time,
                         delta=dt1$censor,marker=dt1$risk.quad.dt1,
                         cause=1,weighting="marginal",
                         times=dt1$time)
ROC.risk.quad.dt1
```

#### Make AUC data

```{r}
# uncensored
uncensored.density1 <- density(dt1$time[dt1$censor==1])
data1_1 <- data.frame(time=uncensored.density1[["x"]], 
                    density=uncensored.density1[["y"]])
data1_1$density.adjusted <- (max(na.omit(ROC.risk.lin.dt1[["AUC"]]))-min(na.omit(ROC.risk.lin.dt1[["AUC"]])))*(uncensored.density1[["y"]]-min(uncensored.density1[["y"]]))/(max(uncensored.density1[["y"]])-min(uncensored.density1[["y"]]))+min(na.omit(ROC.risk.lin.dt1[["AUC"]]))

# censored
censored.density1 <- density(dt1$time[dt1$censor==0])
data1_2 <- data.frame(time=censored.density1[["x"]], 
                    density=censored.density1[["y"]])
data1_2$density.adjusted <- (max(na.omit(ROC.risk.lin.dt1[["AUC"]]))-min(na.omit(ROC.risk.lin.dt1[["AUC"]])))*(censored.density1[["y"]]-min(censored.density1[["y"]]))/(max(censored.density1[["y"]])-min(censored.density1[["y"]]))+min(na.omit(ROC.risk.lin.dt1[["AUC"]]))

res1 <- data.frame(time=ROC.risk.lin.dt1[["times"]],
                    cox.lin.dt1=ROC.risk.lin.dt1[["AUC"]],
                    cox.quad.dt1=ROC.risk.quad.dt1[["AUC"]])
res1 <- melt(data = res1, 
                   id.vars = "time", 
                   measure.vars = c("cox.lin.dt1","cox.quad.dt1"))
```

#### Time dependent AUC plot with ggplot2 (cox.lin.dt1, cox.quad.dt1)

```{r}
dt1$censor1 <- ifelse(dt1$censor==1,
                    1,min(na.omit(ROC.risk.lin.dt1[["AUC"]])))
ggplot() +
  geom_line(data=res1, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt1, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data1_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data1_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(min(dt1$time), max(dt1$time))) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt1, cox.quad.dt1)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.lin.dt1, cox.quad.dt1)

```{r}
mean(ROC.risk.lin.dt1[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.quad.dt1[["AUC"]],na.rm=TRUE) 
```

#### Counting censored data

```{r}
dt1.censor <- dt1[dt1$censor==0,]
dt1.censor <- dt1.censor %>%
  mutate(n.cen=cut(dt1.censor$time, 
                   breaks=seq(from=0,to=1000,by=50)))
table(dt1.censor$n.cen)
```




## Assume quadratic relationship

```{r}
# Assume parametric basis distribution is Exponential distribution... 
lambda1 <- 0.01 # parameter for generate Y  
lambda2 <- 0.01 # parameter for generate C 
beta1 <- 3 ; beta2 <- 2 # true coefficients
Y <- c() # real survival time
C <- c() # censoring time
t <- c() # observed survival time
censor <- c() # censoring indicator
X <- c() # explanatory variable
U1 <- c() # random variable from U(0,1) for generate Y
U2 <- c() # random variable from U(0,1) for generate C 
X2 <- c() # X^2

for (i in 1:1000) {
  U1[i] <- runif(1)
  U2[i] <- runif(1)
  # X[i] <- runif(1,min=0,max=10)
  X[i] <- rnorm(1,mean=0,sd=1)
  X2[i] <- X[i]^2
  Y[i] <- -(1/lambda1)*log(U1[i])*exp(-beta1*X[i]-beta2*X2[i])
  C[i] <- -(1/lambda2)*log(U2[i])
  t[i] <- min(Y[i],C[i])
  censor[i] <- ifelse(Y[i]<=C[i],1,0)
}

mean(1-censor) # censoring rate
dt2 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
head(dt2,20)
# dt2[c(order(dt2$time)),]$censor
```


### Time dependent AUC

### True model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1} x + \beta_{2} x^2$
- True coefficients : $\beta_{1}=3, \beta_{2}=2$ 

### Set linear model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x$

```{r}
cox.lin.dt2 <- coxph(Surv(time,censor) ~ X, data=dt2)
summary(cox.lin.dt2)
risk.lin.dt2 <- predict(object=cox.lin.dt2, newdata=dt2, type="risk") # risk score
dt2$risk.lin.dt2 <- risk.lin.dt2
ROC.risk.lin.dt2 <- timeROC(T=dt2$time,
                         delta=dt2$censor,marker=dt2$risk.lin.dt2,
                         cause=1,weighting="marginal",
                         times=dt2$time)
ROC.risk.lin.dt2
```

### Set quadratic model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x+\beta_{2}x^2$

```{r}
cox.quad.dt2 <- coxph(Surv(time,censor) ~ X+X2, data=dt2)
summary(cox.quad.dt2)
risk.quad.dt2 <- predict(object=cox.quad.dt2, newdata=dt2, type="risk") # risk score
dt2$risk.quad.dt2 <- risk.quad.dt2
ROC.risk.quad.dt2 <- timeROC(T=dt2$time,
                         delta=dt2$censor,marker=dt2$risk.quad.dt2,
                         cause=1,weighting="marginal",
                         times=dt2$time)
ROC.risk.quad.dt2
```

#### Make AUC data

```{r}
# uncensored
uncensored.density2 <- density(dt2$time[dt2$censor==1])
data2_1 <- data.frame(time=uncensored.density2[["x"]], 
                    density=uncensored.density2[["y"]])
data2_1$density.adjusted <- (max(na.omit(ROC.risk.lin.dt2[["AUC"]]))-min(na.omit(ROC.risk.lin.dt2[["AUC"]])))*(uncensored.density2[["y"]]-min(uncensored.density2[["y"]]))/(max(uncensored.density2[["y"]])-min(uncensored.density2[["y"]]))+min(na.omit(ROC.risk.lin.dt2[["AUC"]]))

# censored
censored.density2 <- density(dt2$time[dt2$censor==0])
data2_2 <- data.frame(time=censored.density2[["x"]], 
                    density=censored.density2[["y"]])
data2_2$density.adjusted <- (max(na.omit(ROC.risk.lin.dt2[["AUC"]]))-min(na.omit(ROC.risk.lin.dt2[["AUC"]])))*(censored.density2[["y"]]-min(censored.density2[["y"]]))/(max(censored.density2[["y"]])-min(censored.density2[["y"]]))+min(na.omit(ROC.risk.lin.dt2[["AUC"]]))

res2 <- data.frame(time=ROC.risk.lin.dt2[["times"]],
                    cox.lin.dt2=ROC.risk.lin.dt2[["AUC"]],
                    cox.quad.dt2=ROC.risk.quad.dt2[["AUC"]])
res2 <- melt(data = res2, 
                   id.vars = "time", 
                   measure.vars = c("cox.lin.dt2","cox.quad.dt2"))
```

#### Time dependent AUC plot with ggplot2 (cox.lin.dt2, cox.quad.dt2)

```{r}
dt2$censor1 <- ifelse(dt2$censor==1,
                    1,min(na.omit(ROC.risk.lin.dt2[["AUC"]])))
ggplot() +
  geom_line(data=res2, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt2, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data2_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data2_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(min(dt2$time), max(dt2$time))) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt2, cox.quad.dt2)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.lin.dt2, cox.quad.dt2)

```{r}
mean(ROC.risk.lin.dt2[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.quad.dt2[["AUC"]],na.rm=TRUE) 
```

#### Counting censored data

```{r}
dt2.censor <- dt2[dt2$censor==0,]
dt2.censor <- dt2.censor %>%
  mutate(n.cen=cut(dt2.censor$time, 
                   breaks=seq(from=0,to=1000,by=50)))
table(dt2.censor$n.cen)
```



## Compare two results

```{r}
a <- ggplot()

fig1 <- a +
  geom_line(data=res1, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt1, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data1_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data1_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, 600)) +
  ggtitle("Assume linear relationship") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))

fig2 <- a +
  geom_line(data=res2, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt2, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data2_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data2_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, 600)) +
  ggtitle("Assume quadratic relationship") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(fig1, fig2, nrow=2, ncol=1)
```



## simulation (repeat 100 times)

### Assume linear relationship

```{r}
seed <- c()
lin.AUC <- c()
quad.AUC <- c()

for (i in 1:100) {
  
  # Assume parametric basis distribution is Exponential distribution... 
  lambda1 <- 0.01 # parameter for generate Y  
  lambda2 <- 0.01 # parameter for generate C 
  beta1 <- 3 # true coefficients
  Y <- c() # real survival time
  C <- c() # censoring time
  t <- c() # observed survival time
  censor <- c() # censoring indicator
  X <- c() # explanatory variable
  U1 <- c() # random variable from U(0,1) for generate Y
  U2 <- c() # random variable from U(0,1) for generate C 
  X2 <- c() # X^2
  
  # Generate simulation data (Assume linear relationship)
  seed[i] <- i
  set.seed(i)
  for (j in 1:1000) {
    U1[j] <- runif(1)
    U2[j] <- runif(1)
    X[j] <- rnorm(1,mean=0,sd=1)
    X2[j] <- X[j]^2
    Y[j] <- -(1/lambda1)*log(U1[j])*exp(-beta1*X[j])
    C[j] <- -(1/lambda2)*log(U2[j])
    t[j] <- min(Y[j],C[j])
    censor[j] <- ifelse(Y[j]<=C[j],1,0)
  }
  dt1 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
  
  # Set linear model
  cox.lin.dt1 <- coxph(Surv(time,censor) ~ X, data=dt1)
  risk.lin.dt1 <- predict(object=cox.lin.dt1, newdata=dt1, type="risk") 
  dt1$risk.lin.dt1 <- risk.lin.dt1
  ROC.risk.lin.dt1 <- timeROC(T=dt1$time,
                         delta=dt1$censor,marker=dt1$risk.lin.dt1,
                         cause=1,weighting="marginal",
                         times=dt1$time)
  
  # Set quadratic model
  cox.quad.dt1 <- coxph(Surv(time,censor) ~ X+X2, data=dt1)
  risk.quad.dt1 <- predict(object=cox.quad.dt1, newdata=dt1, type="risk") 
  dt1$risk.quad.dt1 <- risk.quad.dt1
  ROC.risk.quad.dt1 <- timeROC(T=dt1$time,
                         delta=dt1$censor,marker=dt1$risk.quad.dt1,
                         cause=1,weighting="marginal",
                         times=dt1$time)
  
  # Calculate mean value of time dependent AUC
  lin.AUC[i] <- mean(ROC.risk.lin.dt1[["AUC"]],na.rm=TRUE) 
  quad.AUC[i] <- mean(ROC.risk.quad.dt1[["AUC"]],na.rm=TRUE)
  
  # print(i)
  
}

dat1 <- data.frame(seed=seed, lin.AUC=lin.AUC, quad.AUC=quad.AUC)
plot(x=dat1$seed, y=dat1$lin.AUC, xlab="seed", ylab="mean of AUC",
     main="Assume linear relationship", type="l", col="red", ylim=c(0.85,1))
lines(x=dat1$seed, y=dat1$quad.AUC, type="l", col="blue")
legend("bottomleft",c("lin.AUC","quad.AUC"),col=c("red","blue"),lty=c(1,1))
```

### Assume quadratic relationship

```{r}
seed <- c()
lin.AUC <- c()
quad.AUC <- c()

for (i in 1:100) {
  
  # Assume parametric basis distribution is Exponential distribution... 
  lambda1 <- 0.01 # parameter for generate Y  
  lambda2 <- 0.01 # parameter for generate C 
  beta1 <- 3 ; beta2 <- 2 # true coefficients
  Y <- c() # real survival time
  C <- c() # censoring time
  t <- c() # observed survival time
  censor <- c() # censoring indicator
  X <- c() # explanatory variable
  U1 <- c() # random variable from U(0,1) for generate Y
  U2 <- c() # random variable from U(0,1) for generate C 
  X2 <- c() # X^2

  # Generate simulation data (Assume quadratic relationship)
  seed[i] <- i
  set.seed(i)
  for (j in 1:1000) {
    U1[j] <- runif(1)
    U2[j] <- runif(1)
    X[j] <- rnorm(1,mean=0,sd=1)
    X2[j] <- X[j]^2
    Y[j] <- -(1/lambda1)*log(U1[j])*exp(-beta1*X[j]-beta2*X2[j])
    C[j] <- -(1/lambda2)*log(U2[j])
    t[j] <- min(Y[j],C[j])
    censor[j] <- ifelse(Y[j]<=C[j],1,0)
  }
  dt2 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
  
  # Set linear model
  cox.lin.dt2 <- coxph(Surv(time,censor) ~ X, data=dt2)
  risk.lin.dt2 <- predict(object=cox.lin.dt2, newdata=dt2, type="risk") 
  dt2$risk.lin.dt2 <- risk.lin.dt2
  ROC.risk.lin.dt2 <- timeROC(T=dt2$time,
                         delta=dt2$censor,marker=dt2$risk.lin.dt2,
                         cause=1,weighting="marginal",
                         times=dt2$time)
  
  # Set quadratic model
  cox.quad.dt2 <- coxph(Surv(time,censor) ~ X+X2, data=dt2)
  risk.quad.dt2 <- predict(object=cox.quad.dt2, newdata=dt2, type="risk") 
  dt2$risk.quad.dt2 <- risk.quad.dt2
  ROC.risk.quad.dt2 <- timeROC(T=dt2$time,
                         delta=dt2$censor,marker=dt2$risk.quad.dt2,
                         cause=1,weighting="marginal",
                         times=dt2$time)

  # Calculate mean value of time dependent AUC
  lin.AUC[i] <- mean(ROC.risk.lin.dt2[["AUC"]],na.rm=TRUE) 
  quad.AUC[i] <- mean(ROC.risk.quad.dt2[["AUC"]],na.rm=TRUE)
  
  # print(i)
}

dat2 <- data.frame(seed=seed, lin.AUC=lin.AUC, quad.AUC=quad.AUC)
plot(x=dat2$seed, y=dat2$lin.AUC, xlab="seed", ylab="mean of AUC",
     main="Assume quadratic relationship", type="l", col="red", ylim=c(0.75,1))
lines(x=dat2$seed, y=dat2$quad.AUC, type="l", col="blue")
legend("bottomleft",c("lin.AUC","quad.AUC"),col=c("red","blue"),lty=c(1,1))
```



# Set other distributions

## Background

`-` real survival time $Y$ : Exponential distribution
`-` censoring time $C$ : Uniform distribution

- $h(t|x)=h_{0}(t)exp(f(x))$
- $H(t|x)=\int_{0}^{t} h(s|x)ds= \int_{0}^{t} h_{0}(s)exp(f(x))ds=H_{0}(t)exp(f(x))$
- $h(t|x)=\frac {f(t|x)}{S(t|x)}= -\frac {d}{dt} log[S(t|x)]$
- $log[S(t|x)]= -\int_{0}^{t} h(s|x)ds= -H(t|x)= -H_{0}(t)exp(f(x))$
- $S(t|x)=exp[-H_{0}(t)exp(f(x))]$

- In general,...
- $F(t|x)=1-exp[-H_{0}(t) exp(f(x))]$
- -> $F^{-1}(t|x)=H_{0}^{-1}[-log(1-t) exp(-f(x))]$
- By Inverse Transform Sampling, suppose that $U_{1}, U_{2} \overset{i.i.d}{\sim}\ U(0,1)$, then we can generate real survival time $Y$ from Cox Regression Model using this equation.
- $Y=H_{Y0}^{-1}[-log(U_{1})exp(-f(x))]$
- And by assuming non-informative censoring, we can generate censoring time $C$ like this equation. (Set $Y$ and $C$ are independent each other.)
- $C=H_{C0}^{-1}[-log(U_{2})]$
- And set censoring indicator $\delta=I(Y \leq C)$ and observed survival time $T=min(Y,C)$
- $H_{Y0}(t)$ : Basis cumulative hazard function of real survival time $Y$
- $H_{C0}(t)$ : Basis cumulative hazard function of censoring time $C$

- When parametric basis distribution of real survival time $Y$ is Exponential distribution and parametric basis distribution of censoring time $C$ is Uniform distribution ($Y$ and $C$ are independent each other), then ~
- $Y$ : $f_{Y0}(t)=\lambda e^{-\lambda t}, t>0$ -> $S_{Y0}(t)=e^{-\lambda t}$ -> $h_{Y0}(t)=\lambda$
- $C$ : $f_{C0}(t)=\frac {1}{\lambda}, 0<t<\lambda$ -> $S_{C0}(t)=1-\frac {t}{\lambda}$ -> $h_{C0}(t)=\frac {1}{\lambda -t}$
- $\therefore Y=-\frac{1}{\lambda_{1}} log(U_{1}) exp(-f(x))$, $C=\lambda_{2}(1-U_{2})$


# Generating simulation data


## Assume linear relationship

```{r}
# Assume parametric basis distribution is Exponential distribution... 
lambda1 <- 0.001 # parameter for generate Y  
lambda2 <- 1.5/lambda1 # parameter for generate C 
beta1 <- 3 # true coefficients
Y <- c() # real survival time
C <- c() # censoring time
t <- c() # observed survival time
censor <- c() # censoring indicator
X <- c() # explanatory variable
U1 <- c() # random variable from U(0,1) for generate Y
U2 <- c() # random variable from U(0,1) for generate C 
X2 <- c() # X^2

for (i in 1:1000) {
  U1[i] <- runif(1)
  U2[i] <- runif(1)
  # X[i] <- runif(1,min=0,max=10)
  X[i] <- rnorm(1,mean=0,sd=1)
  X2[i] <- X[i]^2
  Y[i] <- -(1/lambda1)*log(U1[i])*exp(-beta1*X[i])
  C[i] <- (lambda2)*(1-U2[i])
  t[i] <- min(Y[i],C[i])
  censor[i] <- ifelse(Y[i]<=C[i],1,0)
}

mean(1-censor) # censoring rate
dt3 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
head(dt3,20)
# dt3[c(order(dt3$time)),]$censor
```


### Time dependent AUC

### True model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1} x$
- True coefficients : $\beta_{1}=3$ 

### Set linear model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x$

```{r}
cox.lin.dt3 <- coxph(Surv(time,censor) ~ X, data=dt3)
summary(cox.lin.dt3)
risk.lin.dt3 <- predict(object=cox.lin.dt3, newdata=dt3, type="risk") # risk score
dt3$risk.lin.dt3 <- risk.lin.dt3
ROC.risk.lin.dt3 <- timeROC(T=dt3$time,
                         delta=dt3$censor,marker=dt3$risk.lin.dt3,
                         cause=1,weighting="marginal",
                         times=dt3$time)
ROC.risk.lin.dt3
```

### Set quadratic model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x+\beta_{2}x^2$

```{r}
cox.quad.dt3 <- coxph(Surv(time,censor) ~ X+X2, data=dt3)
summary(cox.quad.dt3)
risk.quad.dt3 <- predict(object=cox.quad.dt3, newdata=dt3, type="risk") # risk score
dt3$risk.quad.dt3 <- risk.quad.dt3
ROC.risk.quad.dt3 <- timeROC(T=dt3$time,
                         delta=dt3$censor,marker=dt3$risk.quad.dt3,
                         cause=1,weighting="marginal",
                         times=dt3$time)
ROC.risk.quad.dt3
```

#### Make AUC data

```{r}
# uncensored
uncensored.density3 <- density(dt3$time[dt3$censor==1])
data3_1 <- data.frame(time=uncensored.density3[["x"]], 
                    density=uncensored.density3[["y"]])
data3_1$density.adjusted <- (max(na.omit(ROC.risk.lin.dt3[["AUC"]]))-min(na.omit(ROC.risk.lin.dt3[["AUC"]])))*(uncensored.density3[["y"]]-min(uncensored.density3[["y"]]))/(max(uncensored.density3[["y"]])-min(uncensored.density3[["y"]]))+min(na.omit(ROC.risk.lin.dt3[["AUC"]]))

# censored
censored.density3 <- density(dt3$time[dt3$censor==0])
data3_2 <- data.frame(time=censored.density3[["x"]], 
                    density=censored.density3[["y"]])
data3_2$density.adjusted <- (max(na.omit(ROC.risk.lin.dt3[["AUC"]]))-min(na.omit(ROC.risk.lin.dt3[["AUC"]])))*(censored.density3[["y"]]-min(censored.density3[["y"]]))/(max(censored.density3[["y"]])-min(censored.density3[["y"]]))+min(na.omit(ROC.risk.lin.dt3[["AUC"]]))

res3 <- data.frame(time=ROC.risk.lin.dt3[["times"]],
                    cox.lin.dt3=ROC.risk.lin.dt3[["AUC"]],
                    cox.quad.dt3=ROC.risk.quad.dt3[["AUC"]])
res3 <- melt(data = res3, 
                   id.vars = "time", 
                   measure.vars = c("cox.lin.dt3","cox.quad.dt3"))
```

#### Time dependent AUC plot with ggplot2 (cox.lin.dt3, cox.quad.dt3)

```{r}
dt3$censor1 <- ifelse(dt3$censor==1,
                    1,min(na.omit(ROC.risk.lin.dt3[["AUC"]])))
ggplot() +
  geom_line(data=res3, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt3, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data3_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data3_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(min(dt3$time), max(dt3$time))) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt3, cox.quad.dt3)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.lin.dt3, cox.quad.dt3)

```{r}
mean(ROC.risk.lin.dt3[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.quad.dt3[["AUC"]],na.rm=TRUE) 
```

#### Counting censored data

```{r}
dt3.censor <- dt3[dt3$censor==0,]
dt3.censor <- dt3.censor %>%
  mutate(n.cen=cut(dt3.censor$time, 
                   breaks=seq(from=0,to=1500,by=50)))
table(dt3.censor$n.cen)
```



## Assume quadratic relationship

```{r}
# Assume parametric basis distribution is Exponential distribution... 
lambda1 <- 0.001 # parameter for generate Y  
lambda2 <- 1.5/lambda1 # parameter for generate C 
beta1 <- 3 ; beta2 <- 2 # true coefficients
Y <- c() # real survival time
C <- c() # censoring time
t <- c() # observed survival time
censor <- c() # censoring indicator
X <- c() # explanatory variable
U1 <- c() # random variable from U(0,1) for generate Y
U2 <- c() # random variable from U(0,1) for generate C 
X2 <- c() # X^2

for (i in 1:1000) {
  U1[i] <- runif(1)
  U2[i] <- runif(1)
  # X[i] <- runif(1,min=0,max=10)
  X[i] <- rnorm(1,mean=0,sd=1)
  X2[i] <- X[i]^2
  Y[i] <- -(1/lambda1)*log(U1[i])*exp(-beta1*X[i]-beta2*X2[i])
  C[i] <- (lambda2)*(1-U2[i])
  t[i] <- min(Y[i],C[i])
  censor[i] <- ifelse(Y[i]<=C[i],1,0)
}

mean(1-censor) # censoring rate
dt4 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
head(dt4,20)
# dt4[c(order(dt4$time)),]$censor
```


### Time dependent AUC

### True model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1} x$
- True coefficients : $\beta_{1}=3, \beta_{2}=2$ 

### Set linear model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x$

```{r}
cox.lin.dt4 <- coxph(Surv(time,censor) ~ X, data=dt4)
summary(cox.lin.dt4)
risk.lin.dt4 <- predict(object=cox.lin.dt4, newdata=dt4, type="risk") # risk score
dt4$risk.lin.dt4 <- risk.lin.dt4
ROC.risk.lin.dt4 <- timeROC(T=dt4$time,
                         delta=dt4$censor,marker=dt4$risk.lin.dt4,
                         cause=1,weighting="marginal",
                         times=dt4$time)
ROC.risk.lin.dt4
```

### Set quadratic model

- $h(t|x)=h _{0} (t)exp[f(x)]$
- $f(x)=\beta_{1}x+\beta_{2}x^2$

```{r}
cox.quad.dt4 <- coxph(Surv(time,censor) ~ X+X2, data=dt4)
summary(cox.quad.dt4)
risk.quad.dt4 <- predict(object=cox.quad.dt4, newdata=dt4, type="risk") # risk score
dt4$risk.quad.dt4 <- risk.quad.dt4
ROC.risk.quad.dt4 <- timeROC(T=dt4$time,
                         delta=dt4$censor,marker=dt4$risk.quad.dt4,
                         cause=1,weighting="marginal",
                         times=dt4$time)
ROC.risk.quad.dt4
```

#### Make AUC data

```{r}
# uncensored
uncensored.density4 <- density(dt4$time[dt4$censor==1])
data4_1 <- data.frame(time=uncensored.density4[["x"]], 
                    density=uncensored.density4[["y"]])
data4_1$density.adjusted <- (max(na.omit(ROC.risk.lin.dt4[["AUC"]]))-min(na.omit(ROC.risk.lin.dt4[["AUC"]])))*(uncensored.density4[["y"]]-min(uncensored.density4[["y"]]))/(max(uncensored.density4[["y"]])-min(uncensored.density4[["y"]]))+min(na.omit(ROC.risk.lin.dt4[["AUC"]]))

# censored
censored.density4 <- density(dt4$time[dt4$censor==0])
data4_2 <- data.frame(time=censored.density4[["x"]], 
                    density=censored.density4[["y"]])
data4_2$density.adjusted <- (max(na.omit(ROC.risk.lin.dt4[["AUC"]]))-min(na.omit(ROC.risk.lin.dt4[["AUC"]])))*(censored.density4[["y"]]-min(censored.density4[["y"]]))/(max(censored.density4[["y"]])-min(censored.density4[["y"]]))+min(na.omit(ROC.risk.lin.dt4[["AUC"]]))

res4 <- data.frame(time=ROC.risk.lin.dt4[["times"]],
                    cox.lin.dt4=ROC.risk.lin.dt4[["AUC"]],
                    cox.quad.dt4=ROC.risk.quad.dt4[["AUC"]])
res4 <- melt(data = res4, 
                   id.vars = "time", 
                   measure.vars = c("cox.lin.dt4","cox.quad.dt4"))
```

#### Time dependent AUC plot with ggplot2 (cox.lin.dt4, cox.quad.dt4)

```{r}
dt4$censor1 <- ifelse(dt4$censor==1,
                    1,min(na.omit(ROC.risk.lin.dt4[["AUC"]])))
ggplot() +
  geom_line(data=res4, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt4, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data4_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data4_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(min(dt4$time), max(dt4$time))) +
  ggtitle("Time dependent AUC with risk score (cox.lin.dt4, cox.quad.dt4)") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Mean value of Time dependent AUC (cox.lin.dt4, cox.quad.dt4)

```{r}
mean(ROC.risk.lin.dt4[["AUC"]],na.rm=TRUE) 
mean(ROC.risk.quad.dt4[["AUC"]],na.rm=TRUE) 
```

#### Counting censored data

```{r}
dt4.censor <- dt4[dt4$censor==0,]
dt4.censor <- dt4.censor %>%
  mutate(n.cen=cut(dt4.censor$time, 
                   breaks=seq(from=0,to=1500,by=50)))
table(dt4.censor$n.cen)
```



## Compare two results

```{r}
a <- ggplot()

fig1 <- a +
  geom_line(data=res3, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt3, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data3_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data3_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, 1500)) +
  ggtitle("Assume linear relationship") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))

fig2 <- a +
  geom_line(data=res4, aes(x=time, y=value,
                            group=variable, color=variable)) +
  geom_point(data=dt4, aes(x=time, y=censor1, color=as.factor(censor))) +
  geom_line(data=data4_1, aes(x=time, y=density.adjusted), 
            size=1, colour='black') +
  geom_line(data=data4_2, aes(x=time, y=density.adjusted), 
            size=1, colour='orange') +
  scale_x_continuous(limits = c(0, 1500)) +
  ggtitle("Assume quadratic relationship") +
  xlab("time") + ylab("AUC") +
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(fig1, fig2, nrow=2, ncol=1)
```



## simulation (repeat 100 times)

### Assume linear relationship

```{r}
seed <- c()
lin.AUC <- c()
quad.AUC <- c()

for (i in 1:100) {
  
  # Assume parametric basis distribution is Exponential distribution... 
  lambda1 <- 0.001 # parameter for generate Y  
  lambda2 <- 1.5/lambda1 # parameter for generate C 
  beta1 <- 3 # true coefficients
  Y <- c() # real survival time
  C <- c() # censoring time
  t <- c() # observed survival time
  censor <- c() # censoring indicator
  X <- c() # explanatory variable
  U1 <- c() # random variable from U(0,1) for generate Y
  U2 <- c() # random variable from U(0,1) for generate C 
  X2 <- c() # X^2

  # Generate simulation data (Assume linear relationship)
  seed[i] <- i
  set.seed(i)
  for (j in 1:1000) {
    U1[j] <- runif(1)
    U2[j] <- runif(1)
    X[j] <- rnorm(1,mean=0,sd=1)
    X2[j] <- X[j]^2
    Y[j] <- -(1/lambda1)*log(U1[j])*exp(-beta1*X[j])
    C[j] <- (lambda2)*(1-U2[j])
    t[j] <- min(Y[j],C[j])
    censor[j] <- ifelse(Y[j]<=C[j],1,0)
  }
  dt3 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)
  
  # Set linear model
  cox.lin.dt3 <- coxph(Surv(time,censor) ~ X, data=dt3)
  risk.lin.dt3 <- predict(object=cox.lin.dt3, newdata=dt3, type="risk") 
  dt3$risk.lin.dt3 <- risk.lin.dt3
  ROC.risk.lin.dt3 <- timeROC(T=dt3$time,
                         delta=dt3$censor,marker=dt3$risk.lin.dt3,
                         cause=1,weighting="marginal",
                         times=dt3$time)
  
  # Set quadratic model
  cox.quad.dt3 <- coxph(Surv(time,censor) ~ X+X2, data=dt3)
  risk.quad.dt3 <- predict(object=cox.quad.dt3, newdata=dt3, type="risk") 
  dt3$risk.quad.dt3 <- risk.quad.dt3
  ROC.risk.quad.dt3 <- timeROC(T=dt3$time,
                         delta=dt3$censor,marker=dt3$risk.quad.dt3,
                         cause=1,weighting="marginal",
                         times=dt3$time)
  
  # Calculate mean value of time dependent AUC
  lin.AUC[i] <- mean(ROC.risk.lin.dt3[["AUC"]],na.rm=TRUE) 
  quad.AUC[i] <- mean(ROC.risk.quad.dt3[["AUC"]],na.rm=TRUE)
  
  # print(i)
  
}

dat3 <- data.frame(seed=seed, lin.AUC=lin.AUC, quad.AUC=quad.AUC)
plot(x=dat3$seed, y=dat3$lin.AUC, xlab="seed", ylab="mean of AUC",
     main="Assume linear relationship \n (Y : Exponential, C : Uniform)", type="l", col="red", ylim=c(0.85,1))
lines(x=dat3$seed, y=dat3$quad.AUC, type="l", col="blue")
legend("bottomleft",c("lin.AUC","quad.AUC"),col=c("red","blue"),lty=c(1,1))
```

### Assume quadratic relationship

```{r}
seed <- c()
lin.AUC <- c()
quad.AUC <- c()

for (i in 1:100) {
  
  # Assume parametric basis distribution is Exponential distribution... 
  lambda1 <- 0.001 # parameter for generate Y  
  lambda2 <- 1.5/lambda1 # parameter for generate C 
  beta1 <- 3 ; beta2 <- 2 # true coefficients
  Y <- c() # real survival time
  C <- c() # censoring time
  t <- c() # observed survival time
  censor <- c() # censoring indicator
  X <- c() # explanatory variable
  U1 <- c() # random variable from U(0,1) for generate Y
  U2 <- c() # random variable from U(0,1) for generate C 
  X2 <- c() # X^2

  # Generate simulation data (Assume quadratic relationship)
  seed[i] <- i
  set.seed(i)
  for (j in 1:1000) {
    U1[j] <- runif(1)
    U2[j] <- runif(1)
    X[j] <- rnorm(1,mean=0,sd=1)
    X2[j] <- X[j]^2
    Y[j] <- -(1/lambda1)*log(U1[j])*exp(-beta1*X[j]-beta2*X2[j])
    C[j] <- (lambda2)*(1-U2[j])
    t[j] <- min(Y[j],C[j])
    censor[j] <- ifelse(Y[j]<=C[j],1,0)
  }
  dt4 <- data.frame(X=X, X2=X2, Y=Y, C=C, time=t, censor=censor)

  # Set linear model
  cox.lin.dt4 <- coxph(Surv(time,censor) ~ X, data=dt4)
  risk.lin.dt4 <- predict(object=cox.lin.dt4, newdata=dt4, type="risk") 
  dt4$risk.lin.dt4 <- risk.lin.dt4
  ROC.risk.lin.dt4 <- timeROC(T=dt4$time,
                         delta=dt4$censor,marker=dt4$risk.lin.dt4,
                         cause=1,weighting="marginal",
                         times=dt4$time)
  
  # Set quadratic model
  cox.quad.dt4 <- coxph(Surv(time,censor) ~ X+X2, data=dt4)
  risk.quad.dt4 <- predict(object=cox.quad.dt4, newdata=dt4, type="risk") 
  dt4$risk.quad.dt4 <- risk.quad.dt4
  ROC.risk.quad.dt4 <- timeROC(T=dt4$time,
                         delta=dt4$censor,marker=dt4$risk.quad.dt4,
                         cause=1,weighting="marginal",
                         times=dt4$time)

  # Calculate mean value of time dependent AUC
  lin.AUC[i] <- mean(ROC.risk.lin.dt4[["AUC"]],na.rm=TRUE) 
  quad.AUC[i] <- mean(ROC.risk.quad.dt4[["AUC"]],na.rm=TRUE)
  
  # print(i)
  
}

dat4 <- data.frame(seed=seed, lin.AUC=lin.AUC, quad.AUC=quad.AUC)
plot(x=dat4$seed, y=dat4$lin.AUC, xlab="seed", ylab="mean of AUC",
     main="Assume quadratic relationship \n (Y : Exponential, C : Uniform)", type="l", col="red", ylim=c(0.75,1))
lines(x=dat4$seed, y=dat4$quad.AUC, type="l", col="blue")
legend("bottomleft",c("lin.AUC","quad.AUC"),col=c("red","blue"),lty=c(1,1))
```
